---
title: Data from Litter removal in a tropical rain forest reduces fine root biomass
  and production but litter addition has few effects
author: Jeff Walker
date: '2018-06-20'
slug: data-from-litter-removal-in-a-tropical-rain-forest-reduces-fine-root-biomass-and-production-but-litter-addition-has-few-effects
categories: []
tags:
  - linear mixed model
  - ecology
  - experiment
  - factorial
  - model selection
subtitle: ''
---
# Sources

Rodtassana C, Tanner EVJ (2018) Litter removal in a tropical rain forest reduces fine root biomass and production but litter addition has few effects. Ecology 99(3): 735-742. https://doi.org/10.1002/ecy.2143

Rodtassana C, Tanner EVJ (2018) Data from: Litter removal in a tropical rain forest reduces fine root biomass and production but litter addition has few effects. Dryad Digital Repository. https://doi.org/10.5061/dryad.f1948

# Background

## Reproducibility key
1. Reproducibility: 1 (fails to reproduce). Not really close
2. Reproducibility: 2 (reproduces to a similar value). Differences may be due to slightly different specification of model or to slight differences in data (case may be excluded or missing)
3. Reproducibility: 3 (reproduces to decimal or random error). Differences due to rounding or stochastic part of estimation.
4. Reproducibility: 3- (reproduces to something close to random error). Differences very small and due to either different specification or stochastic or maybe algorithmic differences between packages.

A "+" means that the precise method wasn't specified and reproducibility required exploration.

## flags

1. red flag: a method that I think is inferior, or is problematic or is simply wrong, or an interpretation that is wrong, or similar issues with the analysis.
2. code flag: a note on a "best practice" or at least better practice alternative to increase reproducibility related to the code.
3. work flow flag: a note on a "best practice" or at least better practice alternative to increase reproducibility related to the organization of the data and analysis.
4. communication flag: a note on a "best practice" or at least better practice alternative to improve communication of results.

## Caveats
Failure to reproduce (level 1) or partial reproducibility (level 2 or 3-) very well might be an error in my coding, or my misunderstanding of the author's methods, or my lack of knowledge of statistics generally or the R functions that I use more specifically. Similarly, my **Red Flags** may also represent my own lack of knowledge or best practices in statistics.

# Reproducibility

```{r prep, message=FALSE}
library(knitr)
library(ggplot2)
library(readxl)
library(data.table)
library(emmeans)
library(nlme)
library(lme4)
library(lmerTest)

base_path <- "../data" #knitr
folder <- '2018-06-20-data-from-litter-removal-in-a-tropical-rain-forest'
fn <- 'Root_Ecology_supplementary_materials.xlsx'
file_path <- paste(base_path, folder, fn, sep='/')
root <- data.table(read_excel(file_path, range = cell_rows(1:16)))
```

Some code used for model selection with a mixed model following Chapter 5.7 of Zuur for model selection of mixed effect models.

```{r modelSelection}
select_model <- function(y, x, r, dt, reml=FALSE){
  # the data are in dt
  # y is the response column name
  # x is the list of fixed effect column names
  # r is the list of random effect column names
  # it would seem wierd to have both month and plot in model
  # see 5.7 of Zuur for model selection of mixed effect models
  fixed <- paste(y,"~", paste(x, collapse='*'))
  random <- list()
  random[[1]] <- paste(paste("(1|", r, ")", sep=""), collapse=" + ")
  min_aic <- 1
  
  # fit random part first, if there are > 1 random variables
  if(length(r) > 1){
    aic <- numeric(3)
    form <- formula(paste(fixed, random[[1]], sep=' + '))
    fit1 <- lmer(form, data=dt, REML=TRUE)
    aic[1] <- AIC(fit1)
    r1 <- r[1] # just the first random element
    random[[2]] <- paste(paste("(1|", r1, ")", sep=""), collapse=" + ")
    form <- formula(paste(fixed, random[[2]], sep=' + '))
    fit2 <- lmer(form, data=dt, REML=TRUE)
    aic[2] <- AIC(fit2)
    r2 <- r[2] # just the second random element
    random[[3]] <- paste(paste("(1|", r2, ")", sep=""), collapse=" + ")
    form <- formula(paste(fixed, random[[3]], sep=' + '))
    fit3 <- lmer(form, data=dt, REML=TRUE)
    aic[3] <- AIC(fit3)
    min_aic <- which(aic == min(aic))
  }
  random <- random[[min_aic]]

  # fit fixed part
  form <- list()
  # fit with ML, choose model, then refit with REML
  aic <- numeric(4)
  form[[1]] <- formula(paste(fixed, random, sep=' + '))
  # default is REML
  fit1 <- lmer(form[[1]], data=dt, REML=reml)
  aic[1] <- AIC(fit1)
  
  fixed <- paste(y,"~", paste(x, collapse=' + '))
  form[[2]] <- formula(paste(fixed, random, sep=' + '))
  fit2 <- lmer(form[[2]], data=dt, REML=reml)
  aic[2] <- AIC(fit2)

  fixed <- paste(y,"~", x[1])
  form[[3]] <- formula(paste(fixed, random, sep=' + '))
  fit3 <- lmer(form[[3]], data=dt, REML=reml)
  aic[3] <- AIC(fit3)

  fixed <- paste(y,"~", x[2])
  form[[4]] <- formula(paste(fixed, random, sep=' + '))
  fit4 <- lmer(form[[4]], data=dt, REML=reml)
  aic[4] <- AIC(fit4)

  min_aic <- which(aic == min(aic))
  return(form[[min_aic]])
}

```

Note in Excel data sheet
> FRB = Fine root biomass (g/m2) at 0-5 cm and 0-10 cm
FRL = Fine root length (m/m2 of root window) at 0-5 cm and 0-10 cm
Sur = Fine root length survivorship (m/m2 of root window) at 0-10 cm
Pro =Fine root length production (m/m2 of root window) at 0-10 cm
IC = Fine root production from ingrowth cores (g/m2/3-months)

## Fine Root Biomass

```{r FRB}
# root is in wide format - reshape to long
FRB_cols <- c("FRB5cmMar13", "FRB5cmApr13", "FRB5cmMay13", "FRB5cmJun13", "FRB5cmJul13", "FRB5cmAug13", "FRB5cmSep13", "FRB5cmOct13", "FRB5cmNov13", "FRB5cmDec13", "FRB5cmJan14", "FRB5cmFeb14", "FRB10cmMar13", "FRB10cmApr13", "FRB10cmMay13", "FRB10cmJun13", "FRB10cmJul13", "FRB10cmAug13", "FRB10cmSep13", "FRB10cmOct13", "FRB10cmNov13", "FRB10cmDec13")
# what are these "FRB5cmJul13__1", "FRB10cmJul13__1" they have values more like FRL than FRB
frb <- melt(root, id.vars=c('treatment', 'plot'), measure.vars <- FRB_cols, variable.name='label', value.name='FRB')
frb[, treatment:=factor(treatment)]
frb[, label:=as.character(label)]
frb[, depth:=factor(ifelse(substr(label, 4, 4)=='5', 5, 10))]
frb[, month:=substr(label, nchar(label)-4, nchar(label)-2)]
dry <- c('Jan', 'Feb', 'Mar', 'Apr')
frb[, season:=factor(ifelse(month %in% dry, 'dry', 'wet'))]

```

### Figure 1
```{r FRBplot, fig.cap="Reproducing Figure 1"}
frb5 <- frb[depth==5, .(N=.N,
              FRB=mean(FRB),
              SE=sd(FRB)/sqrt(.N)),
              by=.(treatment, season)]
frb10 <- frb[depth==10, .(N=.N,
              FRB=mean(FRB),
              SE=sd(FRB)/sqrt(.N)),
              by=.(treatment, season)]
frb_sum <- rbind(data.table(depth=5, frb5), data.table(depth=10, frb10))
frb_sum[, Lower:=FRB-SE]
frb_sum[, Upper:=FRB+SE]
frb_sum[, season:=factor(season, c('wet', 'dry'))]
frb_sum[, treatment:=factor(treatment, c('LR', 'CT', 'LA'))]
pd <- position_dodge(.6)
gg <- ggplot(data=frb_sum, aes(x=season, y=FRB, fill=treatment)) +
  geom_col(position=pd, width=0.6, color='black') +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2, position=pd) +
  scale_fill_manual(values=c("white", "gray", "black")) +
  facet_grid(.~depth, labeller = label_both) +
  theme_minimal(base_size=14)
gg

```

### 5 cm depth

> Fine root mass in the soil was higher in the wet season than in the dry season, both at 0-5 cm soil depth (F1, 178 = 10.2, P<0.01, Fig. 1)

```{r FRB_5cm}
y <- 'FRB'
x <- c('treatment', 'season')
r <- c('plot', 'month') # should this include "month"?
dt <- frb[depth==5]
# model select
form <- select_model(y, x, r, dt, reml=FALSE)
# anova
anova(lmer(form, data=dt, REML=FALSE), type=3)
# anova(lmer(form, data=dt, REML=TRUE), type=3)
anova(lmer(form, data=dt, REML=FALSE), type=3, ddf="Kenward-Roger")
anova(lmer(form, data=dt, REML=FALSE), type=3, ddf="lme4")

# manual F
1-pf(10.1538, 1, 11.823)
# published numbes
1-pf(10.1538, 1, 178) # hmm. okay so not < 0.001

```

Reproducibility: 2+ (reproduces to a similar value). The F reproduces at least to rounding error only if the final model is fit with REML=FALSE. But the df are very different and they don't give an exact p-value. The authors do not report the structure of the final model, which makes it hard to check.

> Fine root mass was lower in litter removal soils at 0-5 cm (significant, Fig. 1)... Litter addition did not significantly lower fine root mass in either 0-5 cm or 0-10 cm soils (Fig. 1 and Appendix S1: Fig. S1).

```{r FRB5cmContrasts}

fit <- lmer(form, data=dt, REML=FALSE)
b_table <- emmeans(fit, pairwise ~ treatment | season)[[2]]
kable(b_table, digits=3, caption='Tukey')
```

Reproducibility: 3 (reproduces to decimal or random error). The pattern is the same but the authors only report letters indicating significance clusters.

### 10 cm depth

> and at 0-10 cm soil depth (F1, 178 = 14.9, P<0.001, Fig. 1)

```{r FRB_10cm}
y <- 'FRB'
x <- c('treatment', 'season')
r <- c('plot', 'month') # should this include "month"?
dt <- frb[depth==10]
# model select
form <- select_model(y, x, r, dt, reml=FALSE)
# anova
anova(lmer(form, data=dt, REML=FALSE), type=3)

```

This isn't close. Explore
```{r FRB_10cm_explore}
form <- formula(FRB ~ season*treatment + (1|plot) + (1|month))
anova(lmer(form, data=dt, REML=FALSE))

form <- formula(FRB ~ season+treatment + (1|plot) + (1|month))
anova(lmer(form, data=dt, REML=FALSE))

form <- formula(FRB ~ season + (1|plot) + (1|month))
anova(lmer(form, data=dt, REML=FALSE))

form <- formula(FRB ~ season*treatment + (1|plot))
anova(lmer(form, data=dt, REML=FALSE))

form <- formula(FRB ~ season+treatment + (1|plot))
anova(lmer(form, data=dt, REML=FALSE))

form <- formula(FRB ~ season + (1|plot))
anova(lmer(form, data=dt, REML=FALSE))

form <- formula(FRB ~ season*treatment + (1|month))
anova(lmer(form, data=dt, REML=FALSE))

form <- formula(FRB ~ season*treatment)
anova(lm(form, data=dt))

form <- formula(FRB ~ season + treatment)
anova(lm(form, data=dt))

```

Reproducibility: 1 (fails to reproduce). None of these result in F with anything near 14.9

### FRL exploration
```{r FRC_10_explore_more}
form <- formula(FRB ~ season*treatment + (1|plot) + (1|month))
fit <- lmer(form, data=dt, REML=FALSE)
emmeans(fit, pairwise ~ treatment | season)# [[2]]
emmeans(fit, pairwise ~ treatment)[[2]]
form <- formula(FRB ~ treatment + (1|plot) + (1|month))
fit <- lmer(form, data=dt, REML=FALSE)
emmeans(fit, pairwise ~ treatment)[[2]]


```


## Fine Root Length
```{r FRL}
FRL_cols <- c("FRL5cmJun13", "FRL5cmAug13", "FRL5cmSep13", "FRL5cmOct13", "FRL5cmNov13", "FRL5cmDec13", "FRL5cmJan14", "FRL5cmFeb14", "FRL5cmMar14", "FRL5cmApr14", "FRL5cmMay14", "FRL10cmJun13", "FRL10cmAug13", "FRL10cmSep13",  "FRL10cmOct13", "FRL10cmNov13", "FRL10cmDec13", "FRL10cmJan14", "FRL10cmFeb14", "FRL10cmMar14", "FRL10cmApr14", "FRL10cmMay14")
# what are these "FRB5cmJul13__1", "FRB10cmJul13__1" they have values more like FRL than FRB
frl <- melt(root, id.vars=c('treatment', 'plot'), measure.vars <- FRL_cols, variable.name='label', value.name='FRL')
frl[, treatment:=factor(treatment)]
frl[, label:=as.character(label)]
frl[, depth:=factor(ifelse(substr(label, 4, 4)=='5', 5, 10))]
frl[, month:=substr(label, nchar(label)-4, nchar(label)-2)]
dry <- c('Jan', 'Feb', 'Mar', 'Apr')
frl[, season:=factor(ifelse(month %in% dry, 'dry', 'wet'))]

```

### Figure 3 A & B

```{r FRLplot, fig.cap="Reproducing Figure 3 A & B. I've combined into single plot"}
frl5 <- frl[depth==5, .(N=.N,
              FRL=mean(FRL),
              SE=sd(FRL)/sqrt(.N)),
              by=.(treatment, season)]
frl10 <- frl[depth==10, .(N=.N,
              FRL=mean(FRL),
              SE=sd(FRL)/sqrt(.N)),
              by=.(treatment, season)]
frl_sum <- rbind(data.table(depth=5, frl5), data.table(depth=10, frl10))
frl_sum[, Lower:=FRL-SE]
frl_sum[, Upper:=FRL+SE]
frl_sum[, season:=factor(season, c('wet', 'dry'))]
frl_sum[, treatment:=factor(treatment, c('LR', 'CT', 'LA'))]
pd <- position_dodge(.6)
gg <- ggplot(data=frl_sum, aes(x=season, y=FRL, fill=treatment)) +
  geom_col(position=pd, width=0.6, color='black') +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2, position=pd) +
  scale_fill_manual(values=c("white", "gray", "black")) +
  facet_grid(.~depth, labeller = label_both) +
  theme_minimal(base_size=14)
gg

```

### 5 cm depth

```{r FRL_5cm}
y <- 'FRL'
x <- c('treatment', 'season')
r <- c('plot', 'month')
dt <- frl[depth==5]

# model select
form <- select_model(y, x, r, dt, reml=FALSE)
# anova
anova(lmer(form, data=dt, REML=FALSE), type=3)
```

Reproducibility: 2 (reproduces to a similar value). Differences may be due to slightly different specification of model or to slight differences in data (case may be excluded or missing)

> Seasonal changes in rainfall had similar effects on standing fine root length as for fine root mass, which overall was higher in the wet season for 0-5 cm soil (Table 1 and Fig. 3a)

**red flag** Table 1 and Fig. 3a do not provide the inference for this because of the substantial interaction, with the direction of the season effect for control opposite of that for LA and LR.

```{r FRL_season}
fit <- lmer(form, data=dt, REML=FALSE)
emmeans(fit, pairwise ~ season)[[2]]
```
This supports the statement generally but it should be qualified given the interaction.

> There was a significant interaction between treatment and season; standing length in litter removal was not significantly lower than controls in the wet season but was significantly lower in the dry season (Figs. 3a and 3b). 

```{r FRL5Contrasts}
fit <- lmer(form, data=dt, REML=FALSE)
b_table <- emmeans(fit, pairwise ~ treatment | season)[[2]]
kable(b_table, digits=3, caption='Tukey')

```

Reproducibility: 3 (reproduces to decimal or random error). The pattern is the same but the authors only report letters indicating significance clusters.

**communications flag** -- "not significantly lower than controls in the wet season" is a strange way to report this since the LA treatment is *greater than* control. It's not simply the same pattern but "not statistically significant"

### 10 cm depth

```{r FRL_10cm}
y <- 'FRL'
x <- c('treatment', 'season')
r <- c('plot', 'month')
dt <- frl[depth==10]

# model select
form <- select_model(y, x, r, dt, reml=FALSE)
# anova
anova(lmer(form, data=dt, REML=FALSE), type=3)
```

Reproducibility: 2 (reproduces to a similar value). Differences may be due to slightly different specification of model or to slight differences in data (case may be excluded or missing)

**communication flag** the authors should report the exact p-value instead of "n.s." as (1) n.s. is not reproducible and (2) readers may have different interpretations of p (not NHST).

```{r FRL10Contrast}
fit <- lmer(form, data=dt, REML=FALSE)
b_table <- emmeans(fit, pairwise ~ treatment | season)[[2]]
kable(b_table, digits=3, caption='Tukey')

```

Reproducibility: 3 (reproduces to decimal or random error). The pattern is the same but the authors only report letters indicating significance clusters.

**communications flag** -- Same as for 5 cm, that is, "not significantly lower than controls in the wet season" is a strange way to report this since the LA treatment is *greater than* control. It's not simply the same pattern but "not statistically significant"

## Root Length Production

```{r RLP}
rlp_cols <- c("ProJul13", "ProAug13", "ProSep13", "ProOct13", "ProNov13", "ProDec13", "ProJan14", "ProFeb14", "ProMar14", "ProApr14", "ProMay14")

rlp <- melt(root, id.vars=c('treatment', 'plot'), measure.vars <- rlp_cols, variable.name='label', value.name='Prod')
rlp[, treatment:=factor(treatment)]
rlp[, label:=as.character(label)]
rlp[, depth:=factor(ifelse(substr(label, 4, 4)=='5', 5, 10))]
rlp[, month:=substr(label, nchar(label)-4, nchar(label)-2)]
dry <- c('Jan', 'Feb', 'Mar', 'Apr')
rlp[, season:=factor(ifelse(month %in% dry, 'dry', 'wet'))]

# sample size
# rlp[, .(N=.N), by=.(treatment, season, depth)]
# rlp[, .(N=.N), by=.(treatment, season, depth, month)]
# rlp[, .(N=.N), by=.(treatment, season, plot)]
```

### Figure 3C

```{r fig3c}
rlp_sum <- rlp[,.(N=.N,
              Prod=mean(Prod),
              SE=sd(Prod)/sqrt(.N)),
              by=.(treatment, season)]
rlp_sum[, Lower:=Prod-SE]
rlp_sum[, Upper:=Prod+SE]
rlp_sum[, season:=factor(season, c('wet', 'dry'))]
rlp_sum[, treatment:=factor(treatment, c('LR', 'CT', 'LA'))]
pd <- position_dodge(.6)
gg <- ggplot(data=rlp_sum, aes(x=season, y=Prod, fill=treatment)) +
  geom_col(position=pd, width=0.6, color='black') +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2, position=pd) +
  scale_fill_manual(values=c("white", "gray", "black")) +
  theme_minimal(base_size=14)
gg

```

### Table 1C (anova)
```{r rlpANOVA}
y <- 'Prod'
x <- c('treatment', 'season')
r <- c('plot', 'month')
dt <- rlp # there is no measure at 5 cm

# model select
form <- select_model(y, x, r, dt, reml=FALSE)
# anova
anova(lmer(form, data=dt, REML=FALSE), type=3)

```

Reproducibility: 1 (fails to reproduce) to 2 (reproduces to a similar value). My F-value for the interaction is massive relative to the reported value. The three values have an underlying similar pattern but the numbers aren't really close.

## Root length survivorship

```{r RLS}
rls_cols <- c("SurJul13", "SurAug13", "SurSep13", "SurOct13", "SurNov13", "SurDec13", "SurJan14", "SurFeb14", "SurMar14", "SurApr14", "SurMay14" )

rls <- melt(root, id.vars=c('treatment', 'plot'), measure.vars <- rls_cols, variable.name='label', value.name='Surv')
rls[, treatment:=factor(treatment)]
rls[, label:=as.character(label)]
rls[, depth:=factor(ifelse(substr(label, 4, 4)=='5', 5, 10))]
rls[, month:=substr(label, nchar(label)-4, nchar(label)-2)]
dry <- c('Jan', 'Feb', 'Mar', 'Apr')
rls[, season:=factor(ifelse(month %in% dry, 'dry', 'wet'))]

```

### Figure 3D

```{r fig3d}
rls_sum <- rls[,.(N=.N,
              Surv=mean(Surv),
              SE=sd(Surv)/sqrt(.N)),
              by=.(treatment, season)]
rls_sum[, Lower:=Surv-SE]
rls_sum[, Upper:=Surv+SE]
rls_sum[, season:=factor(season, c('wet', 'dry'))]
rls_sum[, treatment:=factor(treatment, c('LR', 'CT', 'LA'))]
pd <- position_dodge(.6)
gg <- ggplot(data=rls_sum, aes(x=season, y=Surv, fill=treatment)) +
  geom_col(position=pd, width=0.6, color='black') +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2, position=pd) +
  scale_fill_manual(values=c("white", "gray", "black")) +
  theme_minimal(base_size=14)
gg

```

### Table 1D (anova)

```{r}
y <- 'Surv'
x <- c('treatment', 'season')
r <- c('plot', 'month')
dt <- rls # there is no measure at 5 cm

# model select
form <- select_model(y, x, r, dt, reml=FALSE)
# anova
anova(lmer(form, data=dt, REML=FALSE), type=3)
```

Reproducibility: 2 (reproduces to a similar value). Differences may be due to slightly different specification of model or to slight differences in data (case may be excluded or missing).

