---
title: 'Data from: The significance of prey avoidance behaviour for the maintenance
  of a predator colour polymorphism'
author: Jeff Walker
date: '2019-02-02'
slug: data-from-the-significance-of-prey-avoidance-behaviour-for-the-maintenance-of-a-predator-colour-polymorphism
categories: []
tags:
  - generalized linear mixed model
  - binomial response
subtitle: ''
---

# Sources

Ajuria Ibarra H, Kinahan M, Marcetteau J, Mehigan AJR, Ziegelmeier RO, Reader T (2018) The significance of prey avoidance behaviour for the maintenance of a predator colour polymorphism. Behavioral Ecology, online in advance of print. https://doi.org/10.1093/beheco/ary129

Ajuria Ibarra H, Kinahan M, Marcetteau J, Mehigan AJR, Ziegelmeier RO, Reader T (2018) Data from: The significance of prey avoidance behaviour for the maintenance of a predator colour polymorphism. Dryad Digital Repository. https://doi.org/10.5061/dryad.792r3k1

# Summary

# Setup
```{r warning=FALSE, message=FALSE}
library(ggplot2)
library(readxl)
library(data.table)
library(lme4)
library(lmerTest)
library(emmeans)
library(cowplot)
```


```{r}
data_folder <- "../data"
folder <- "Data from The significance of prey avoidance behaviour for the maintenance of a predator colour polymorphism"

# images need to be moved to /static/images/data from.../image.png
knitr_path <- FALSE # run = TRUE to debug

source("../R/clean_label.R") # bookdown
```

# Import
```{r}
fn <- "Data from prey avoidance and predator colour polymorphism.xlsx"
file_path <- paste(data_folder, folder, fn, sep="/")
sheet_1 <- "Experiment 1"
exp1 <- data.table(read_excel(file_path, sheet=sheet_1))
exp1[, Flower:=factor(Flower)]
exp1[, Spider:=factor(Spider)]

sheet_2 <- "Experiment 2"
exp2 <- data.table(read_excel(file_path, sheet=sheet_2))
colnames(exp2) <- clean_label(colnames(exp2))
exp2[, Treatment_1:=factor(Treatment_1)]
exp2[, Treatment_2:=factor(Treatment_2)]
exp2[, Day_fac:=factor(Day)] # create a new column with day as factor

```


# Reproducibility
## Experiment 1
### Statistical methods
> Treatments were performed in temporal blocks, with each combination offlower species and treatment being applied once in each block. Spiders and treatment-flower combinations were chosen randomly within blocks.

>Honeybee responses in Experiment 1 were analyzed with generalized linear mixed models (GLMMs) with binomial errors and block as a random effect. Binomial responses were: 1) accepted versus not accepted (reject + ignore), 2) accepted versus rejected (with ignored flowers excluded), and 3) inspected (accept + reject) versus not inspected (ignore). Differences between spider treatments were analyzed by comparing a model containing all 4 treatments to a model where the 3 spider treatments were collapsed into one and contrasted with the control.

### cleaning and recoding

1. There is no column identifying block; I will assume that the data are organized into the blocks.

2. The key for the codes in the column "Response" is not given. The responses are "ignored", "reject", "accept" and the codes are 1, 2, 3. It took some trial-and-error to match the code with the response, which must be 1=accept, 2=reject, 3, = ignore

2a. Response[952] has a code value of 13. trial-and-error recoding this to 1 reproduced exactly two of three F values in the first row of Table 1.

3. I attempted to reproduce using three binomial models, one for each response category. This required recoding the "response" column to three separate columns.

4. The column Spider also needed to be recoded to binary (control, not control) for the ANOVA.

```{r recode exp1}
n_combos <- length(levels(exp1$Flower)) * length(levels(exp1$Spider))
n_blocks <- nrow(exp1)/n_combos
exp1[, block:=rep(1:n_blocks, each=n_combos)]

# trial-and-error for item 952
inc <- which(exp1$Response==13)
exp1[inc, Response:=1] 

exp1[, R1:=ifelse(Response==1, 1, 0)] # 1= accept, 0 = not accept
exp1[, R2:=ifelse(Response==1, 1, 
                  ifelse(Response==2, 0, NA))] # 1= accept, 0 = reject, NA = ignore
exp1[, R3:=ifelse(Response==3, 0, 1)] # 1 = inspect (reject or accept), 0 = ignore


exp1[, Spider_binary:=ifelse(Spider=="control", "control", "spider")]
```

```{r trial-and-error, echo=FALSE, eval=FALSE}
exp1[inc, R1:=1] 
m1.R1 <- glmer(R1 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m1.R1, test="Chisq")
m2.R1 <- glmer(R1 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m2.R1, test="Chisq")

exp1[inc, R1:=0] 
m1.R1 <- glmer(R1 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m1.R1, test="Chisq")
m2.R1 <- glmer(R1 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m2.R1, test="Chisq")

exp1[inc, R1:=NA] 
m1.R1 <- glmer(R1 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m1.R1, test="Chisq")
m2.R1 <- glmer(R1 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m2.R1, test="Chisq")

```

### Table 1

```{r}

m1.R1 <- glmer(R1 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
(test1a <- drop1(m1.R1, test="Chisq"))
m2.R1 <- glmer(R1 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
(test1b <- drop1(m2.R1, test="Chisq"))

m1.R2 <- glmer(R2 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
(test2a <- drop1(m1.R2, test="Chisq"))
m2.R2 <- glmer(R2 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
(test2b <- drop1(m2.R2, test="Chisq"))

m1.R3 <- glmer(R3 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
(test3a <- drop1(m1.R3, test="Chisq"))
m2.R3 <- glmer(R3 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
(test3b <- drop1(m2.R3, test="Chisq"))

```

### reported table 1
```{r table1, echo=FALSE}
fn <- "table1.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

#### my table 1

```{r mytable1, echo=FALSE}
spider <- c(test1b["Spider_binary", "LRT"], test2b["Spider_binary", "LRT"], test3b["Spider_binary", "LRT"])
flower <- c(test1b["Flower", "LRT"], test2b["Flower", "LRT"], test3b["Flower", "LRT"])
inter <- c(test1a["Flower:Spider_binary", "LRT"], test2a["Flower:Spider_binary", "LRT"], test3a["Flower:Spider_binary", "LRT"])
responses <- c("Accepted v not accepted","Accepted v rejected", "Inspected v not inspected")
table1 <- data.table(Response=responses,
                     Spider=spider,
                     Flower=flower,
                     Interaction=inter)
knitr::kable(table1, digits=c(NA, 3, 3, 3))
```

Effectively reproduces. The first two numbers in row one are exact. Differences could be to different algorthms between versions. I don't think they are due to differences in coding.

### Effect of spider treatments

>No significant differences were found among spider treatments: the effect of collapsing the spider treatments together in the statistical model was not significant for any of the response variables: accepted versus not accepted (χ2 = 2.718, df = 6, P = 0.843), accepted versus rejected (χ2 = 3.477, df = 6, P = 0.747), and inspected versus not inspected (χ2 = 6.484, df = 6, P = 0.371).

```{r}
m3.R1 <- glmer(R1 ~ Flower*Spider + (1|block), family=binomial, data=exp1)
anova(m1.R1, m3.R1, test="Chisq") # this reproduces

m3.R2 <- glmer(R2 ~ Flower*Spider + (1|block), family=binomial, data=exp1)
anova(m1.R2, m3.R2, test="Chisq")

m3.R3 <- glmer(R3 ~ Flower*Spider + (1|block), family=binomial, data=exp1)
anova(m1.R3, m3.R3, test="Chisq")
```

"accepted versus not accepted" reproduces exactly; the other two are close

### Figure 1

```{r}
m3_R1.emm <- emmeans(m3.R1, specs=c("Flower", "Spider"), type="response")
m3_R2.emm <- emmeans(m3.R2, specs=c("Flower", "Spider"), type="response")
m3_R3.emm <- emmeans(m3.R3, specs=c("Flower", "Spider"), type="response")

pd <- position_dodge(0.9)
gg1 <- ggplot(data=summary(m3_R1.emm), aes(x=Flower, y=prob, fill=Spider)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax=asymp.UCL), position=pd, width=0.2) +
  coord_cartesian(ylim=c(0, 0.8)) +
  ggtitle("Accept") +
  NULL
gg2 <- ggplot(data=summary(m3_R2.emm), aes(x=Flower, y=prob, fill=Spider)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax=asymp.UCL), position=pd, width=0.2) +
  coord_cartesian(ylim=c(0, 0.8)) +
  ggtitle("Accept after inspection") +
  NULL
gg3 <- ggplot(data=summary(m3_R3.emm), aes(x=Flower, y=prob, fill=Spider)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax=asymp.UCL), position=pd, width=0.2) +
  coord_cartesian(ylim=c(0, 0.8)) +
  ggtitle("Inspect") +
  NULL

# See myfig1 chunk for plot
```

```{r fig1, echo=FALSE}
fn <- "fig1.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

```{r myfig1, fig.asp = 1.5, echo=FALSE}
plot_grid(gg1, gg2, gg3, nrow=3)
```

The CIs differ. My CIs are computed from the model fit using emmeans. These are much more uniform than the reported CIs. The reported CIs may be from confint(fit_model)

## Experiment 2

### Statistical Methods
> For Experiment 2, the differences in number of rejections of flowers by honeybees in Offer 1 (maximum = 10) between those harboring red and yellow spiders, and among the 4 days on which trials took place, were analyzed using a non-parametric 2-way analysis of variance. Responses to spider treatments in Offer 2 were analyzed using generalized linear models (GLMs) with binomial errors. Three binary response variables were used: 1) accepted (fed + land) versus not accepted, 2) fed versus not fed, and 3) whether or not the number of presentations which were rejected in Offer 2 was higher than in Offer 1. For the first 2 of these variables, 2 separate models were fitted, one considering only the  rst attempt to present a flower to a honeybee in Offer 2, and the other considering all attempts to present the flower (maximum = 10; i.e., did the bee ever land/feed?). We expected that any effect of learned aversion to spiders encountered in Offer 1 would be strongest in the first of these 2 analyses. Day of study (1–4) was included as a fixed factor to account for changes in honeybee behavior over time. In all models, the signi cance of each term was assessed using a chi-squared test statistic after backward deletion from a saturated model. A sep- arate chi-squared test was used to assess the difference in the pro- portion of bee responses in each rejection speed class (slow, fast and very fast) among spider treatments for the first presentation of flowers in Offer 2 in Experiment 2. 

### Offer 1 -- nonparametric 2-way ANOVA

> "no significant overall effect of spider color...(2-way nonparametric Anova: H = 0.40, df = 1 P = 0.529) (Figure 2)"...
> "no significant effect of the day of study" (H = 3.43, df = 3, P = 0.331)
> the interaction between spider color and day of study...was not significant (H = 7.46, df = 3, P = 0.060)

Notes on analysis
1. The reported df for $Day$ indicates the authors coded Day as a factor and not numeric. 

2. The authors give no detail on *which* non-parametric 2-way ANOVA was used, so I used a trial-and-error with several R packages.

3. The hint of which non-parametric test used is the report of the testic statistic as "H".

```{r nonparametric libraries, message=FALSE}
library(WRS2)
library(Rfit)
library(ART)
library(ARTool)
library(rcompanion)
```

```{r}
#Rfit
raov(Rejects_in_offer_1 ~ Day_fac*Treatment_1, data=exp2)

```

```{r}
aligned.rank.transform(Rejects_in_offer_1 ~ Day_fac*Treatment_1, data=exp2, SS.type="III")$significance

```

```{r}
t2way(Rejects_in_offer_1 ~ Day_fac*Treatment_1, data=exp2)
pbad2way(Rejects_in_offer_1 ~ Day_fac*Treatment_1, data=exp2)
```

```{r}
scheirerRayHare(Rejects_in_offer_1 ~ Day_fac*Treatment_1, data=exp2)
scheirerRayHare(Rejects_in_offer_1 ~ Treatment_1*Day_fac, data=exp2)
```

```{r exp2-offer1-table, echo=FALSE}
my_H <- scheirerRayHare(Rejects_in_offer_1 ~ Treatment_1*Day_fac, data=exp2, verbose=FALSE)$H
my_p <- scheirerRayHare(Rejects_in_offer_1 ~ Treatment_1*Day_fac, data=exp2, verbose=FALSE)$p.value
reported_H <- c(0.4, 3.43, 7.46)
reported_p <- c(.529, .331, .060)
knitr::kable(data.table("Reported H"=reported_H,
                        "Reported p"=reported_p,
                        "my H"=my_H[1:3],
                        "my p"=my_p[1:3]
                        ), digits=c(3,3,3,3))
```

The Scheirer-Ray-Hare Test has an H statistic and is qualitatively similar to the reported results but does not reproduce quantitatively.

The coding of $Day$, and the interpretation of the test statistic, in combination with the plot, raises Statistical Red Flat no. 1.

### Figure 2

There is no indication on how the SE was computed -- are these raw SE or some robust SE given the non-parametric hypothesis test? They look close-to but not quite uniform, as if they were computed from a model.

```{r}
# raw SE
# there are no NA so .N is accurate
exp2_sum_a <- exp2[, .(mean=mean(Rejects_in_offer_1),
                       SE=sd(Rejects_in_offer_1)/sqrt(.N)), by=.(Day_fac, Treatment_1)]
pd <- position_dodge(0.9)
ggplot(data=exp2_sum_a, aes(x=Day_fac, y=mean, fill=Treatment_1)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=mean-SE, ymax=mean+SE),
                position=pd,
                width=.2) +
  NULL
```

```{r fig2, echo=FALSE}
fn <- "fig2.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

This seems to reproduce (the raw SE). The figure itself is addressed in Statistical Red Flag #1. 

# Statistical Red Flags