---
title: 'Data from: Perceived threats of infanticide reduce maternal allocation during
  lactation and lead to elevated oxidative damage in offspring'
author: Jeff Walker
date: '2019-02-18'
slug: data-from-perceived-threats-of-infanticide-reduce-maternal-allocation-during-lactation-and-lead-to-elevated-oxidative-damage-in-offspring
categories: []
tags:
  - linear mixed model
  - repeated measures anova
  - SPSS
subtitle: ''
output:
  blogdown::html_page:
    toc: true
---

# Sources

Gale T, Garratt M, Brooks RC (2018) Perceived threats of infanticide reduce maternal allocation during lactation and lead to elevated oxidative damage in offspring. Functional Ecology 32(9): 2158-2169. https://doi.org/10.1111/1365-2435.13146

Gale T, Garratt M, Brooks RC (2018) Data from: Perceived threats of infanticide reduce maternal allocation during lactation and lead to elevated oxidative damage in offspring. Dryad Digital Repository. https://doi.org/10.5061/dryad.762k24f

# Summary

1. The available data are clean. All data sets are in a single Excel sheet. An issue for reproducibility is that the data set for each experiment is not identified.
2. In general, most results (that I was able to run) reproduce
3. One error is in the reporting of the ANOVA statistics for growth rate. Two of the terms are reported with the F and P values for other terms in the table.
4. The ANOVA for the interaction effects of Sex x Treatment on weight at adulthood do not reproduce.
5. Oxidative stress cannot be reproduced because the data are incomplete (Two columns, Sex and Plate, were not given)

# Setup
```{r libraries, message=FALSE}
library(ggplot2)
library(readxl)
library(data.table)
library(emmeans)
library(nlme)
library(lme4)
library(lmerTest)
library(afex)
library(car)
```

```{r}
data_folder <- "../data"
folder <- "Data from Perceived threats of infanticide reduce maternal allocation during lactation and lead to elevated oxidative damage in offspring"

# images need to be moved to /static/images/data from.../image.png
knitr_path <- FALSE # run = TRUE to debug

source("../R/clean_label.R") # bookdown
```

# Import
The single sheet contains all experiments in blocks glued wide
```{r}
fn <- "LBE paper data.xlsx"
file_path <- paste(data_folder, folder, fn, sep="/")

# offspring weight
exp1 <- data.table(read_excel(file_path, range="A1:E185"))
colnames(exp1) <- clean_label(colnames(exp1))
exp1[, Treatment:=factor(Treatment)]

exp2 <- data.table(read_excel(file_path, range="F1:H45"))
colnames(exp2) <- clean_label(colnames(exp2))
exp2[, Treatment:=factor(Treatment)]

# maternal behavior
exp3 <- data.table(read_excel(file_path, range="K1:U45"))
colnames(exp3) <- clean_label(colnames(exp3))
exp3[, Treatment:=factor(Treatment)]

# GCMs
exp4 <- data.table(read_excel(file_path, range="V1:Z60"))
colnames(exp4) <- clean_label(colnames(exp4))
exp4[, Treatment:=factor(Treatment)]

#oxidative stress
exp5 <- data.table(read_excel(file_path, range="AA1:AJ86"))
colnames(exp5) <- clean_label(colnames(exp5))
exp5[, Treatment:=factor(Treatment)]

```

# Reproducibility
## Offspring weight

>The mean birth weight of novel male treatment pups (1.619± 0.432g) was not significantly different from that of the familiar male treatment (1.628 ± 0.433g; ANOVA: weight: F1, 182= 0.022, p= 0.882; litter size: F1,42= 0.070, p= 0.792).

```{r}
# birth weight
exp1[,.(birth=mean(birth),
        SD=sd(birth),
        SE=sd(birth)/sqrt(.N)), by=Treatment]
anova(lm(birth ~ Treatment, data=exp1))

# litter size
anova(lm(Littersize ~ Treatment, data=exp2))

```
Reproduces, the error is sample SD and not SEM.

and identifies LBE = novel male treatment, MC = familiar male treatment

>To analyse the differences in offspring growth rate we used repeated measures ANOVA reporting within-subjects effects. Mauchly's Test of Sphericity indicated that the assumption of sphericity had been violated (χ2(2) = 34.210, p = <0.001), therefore degrees of freedom were corrected using Greenhouse-Geisser adjusted degrees of freedom. There was a difference in growth rate between the treatments but there was no effect of sex (treatment: F1.7,310=7.49, p= 0.001; interaction between treatment and time: F1.7,310=5668, p= <0.001; sex x treatment: F1.6,306=1.3, p= 0.270).

A treatment effect on growth rate is indirectly indicated by treatment x time interaction.

```{r}
exp1[, ID:=factor(1:nrow(exp1))]
exp1[, Treatment:=factor(Treatment, c("MC", "LBE"))]
exp1[, Sex:=factor(Sex)]
exp1_long <- melt(exp1, id.vars = c("Treatment", "Sex", "ID"), variable.name="Time", value.name = "Weight")
exp1_long[, ID:=factor(ID)]
#(fit.aov_car <- aov_car(Weight ~ Treatment*Sex*Time + Error(ID/Time), data=exp1_long))
#(fit.ez <- aov_ez("ID", "Weight", exp1_long, between=c("Treatment", "Sex"), within="Time"))
(fit.aov_4 <- aov_4(Weight ~ Treatment*Sex*Time + (Time|ID), data=exp1_long))
```

I think it generally reproduces but I think the authors have mixed up the terms in their reporting.

1. The authors report the treatment x time interaction as F1.7,310=5668, p= <0.001, which 1) is a huge F given what the interaction effect looks like and is nearly equivalent to my stats for the Time term.
2. The authors report sex x treatment: F1.6,306=1.3, p= 0.270 which is very close to my sex x time term.

The unbalanced design raises statistical red flag #1

>This resulted in novel male treatment pups being significantly smaller (7.19±0.61g), at weaning (4 weeks old) (ANOVA: F1,182=27.11, p=<0.001) than the familiar male treatment pups (7.99±1.48g)

```{r}
m1 <- lm(weaning ~ Treatment, data=exp1)
anova(m1)
exp1[, .(weaning=mean(weaning),
         SD=sd(weaning),
         SE=sd(weaning)/sqrt(.N)),
     by=.(Treatment)]
```
The means reproduce but not the sample SD.

>there were no weight differences at adulthood (13 weeks old) (ANOVA: F1, 184=0.760, p=0.384)

```{r}
anova(lm(adulthood ~ Treatment, data=exp1))
```

reproduces

>There were also no significant interactions of treatment and sex in weight at weaning (sex x treatment F1,180= 2.14, p=0.145) or at adulthood (sex x treatment F1, 180=0.044, p=0.868)).

```{r}
anova(lm(weaning ~ Sex*Treatment, data=exp1))
anova(lm(adulthood ~ Sex*Treatment, data=exp1))
```

The weaning results reproduce. The adulthood results do not. Here is a plot of the interaction.

```{r}
(fit.emm <- data.table(summary(emmeans(lm(adulthood ~ Sex*Treatment, data=exp1),
                   specs=c("Sex", "Treatment")))))
setnames(fit.emm, old="emmean", new="adulthood")
pd <- position_dodge(0.5)
gg <- ggplot(data=exp1, aes(x=Treatment, y=adulthood, color=Sex, group=Sex)) +
  #geom_point(position=pd) +
  geom_point(data=fit.emm, aes(y=adulthood), position=pd)+
  geom_errorbar(data=fit.emm, aes(ymin=lower.CL, ymax=upper.CL), position=pd, width=0.2) +
  geom_line(data=fit.emm, position=pd) +
  NULL
gg

```

## Maternal behaviour

>There was no significant difference in the amount of time the mothers spent in nest, grooming, licking pups, eating/ drinking, nest building, resting or active (Table 1). The only behaviour showing significant differences between treatments was nursing, with females exposed to familiar males spent almost twice the time nursing (Table 1).

```{r}
y_list <- c("In_Nest", "Nursing", "Licking_Pups", "Nest_Building", "Eating.Drinking", "Grooming", "Active", "Resting")
my_table <- data.table(NULL)
exp3[, one:=1] # used for getting N in each level
exp3[, myID:=factor(1:.N)] # ID 164 is present twice
for(y in y_list){
  n.mc <- sum(exp3[Treatment=="MC" & !is.na(get(y)), one])
  n.lbe <- sum(exp3[Treatment=="LBE" & !is.na(get(y)), one])
  fixed_i <- formula(paste(y, "Treatment", sep=" ~ "))
  my_table <- rbind(my_table, 
                 data.table(Response=y, 
                            anova(lme(fixed_i, random = ~ 1|Block,
                                      method="ML", data=exp3))[2,],
                            Mean.MC=mean(exp3[Treatment=="MC", get(y)]),
                            SE.MC=sd(exp3[Treatment=="MC", get(y)])/sqrt(n.mc),
                            Mean.LBE=mean(exp3[Treatment=="LBE", get(y)]),
                            SE.LBE=sd(exp3[Treatment=="LBE", get(y)])/sqrt(n.lbe)
                 ))
}
# table is printed below the reported table
```

```{r table1, echo=FALSE}
fn <- "table1.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)
```

```{r mytable, echo=FALSE}
knitr::kable(my_table, digits=c(NA,0,0,2,3,2,2,2,2))
```

Table 1 reproduces with very small differences probably due to implementation of ANOVA/LMM in different statistical software.

## Glucocorticoid metabolites in mothers
> sample sizes for the glucocorticoid metabolites are novel male n=20, familiar male n= 21 and unmated control n= 18. Using a repeated measures ANOVA we found that FGM’s did change significantly over time (effect of time: F2, 112 = 131, p= <0.001; interaction between time and housing companion: F4,112= 29.5, p<0.001).

```{r}
# melt wide to long
exp4_long <- melt(exp4, id.vars=c("ID", "Treatment"), variable.name="Time", value.name="GCM")
(fit.aov_4 <- aov_4(GCM ~ Treatment*Time + (Time|ID), data=exp4_long, anova_table = list(correction = "none")))
```

reproduces

>The was no difference in the FGMs in the females before the experiment began (ANOVA: F2,56=0.094, p= 0.911), but there was following exposure to their partners (ANOVA: F2,56=43.72, p=<0.001) and at the end of the experiment which coincided with mid lactation for the reproducing treatments (F2,56=4.6, p=0.014)(See Figure 3).

```{r}
anova(lm(Collection_1 ~ Treatment, data=exp4))
anova(lm(Collection_2 ~ Treatment, data=exp4))
anova(lm(Collection_3 ~ Treatment, data=exp4))

```

Reproduces

> After removing the unmated control from the analysis we found that the effect of time was still significant (F2, 78= 143, p=<0.001) but the interaction between time and housing companion was not (F1,78=1.89, p=0.158). 

```{r}
aov_4(GCM ~ Treatment*Time + (Time|ID), data=exp4_long[Treatment!="UC",], anova_table = list(correction = "none"))

```

Reproduces

## Fig. 3

```{r}
raw <- exp4_long[!is.na(GCM), .(GCM=mean(GCM),
                     SE=sd(GCM)/sqrt(.N)), by=.(Treatment, Time)]
modeled <- summary(emmeans(fit.aov_4, specs=c("Treatment", "Time")))

pd <- position_dodge(0.5)
gg1 <- ggplot(data=raw, aes(x=Treatment, y=GCM, color=Time, group=Time)) +
  geom_point(position=pd) +
  geom_errorbar(aes(ymin=GCM-SE, ymax=GCM+SE), width=0.2, position=pd) +
  coord_cartesian(ylim=c(0,15000)) + # this is way too much white space
  NULL
gg1

gg1 <- ggplot(data=modeled, aes(x=Treatment, y=emmean, color=Time, group=Time)) +
  geom_point(position=pd) +
  geom_errorbar(aes(ymin=emmean-SE, ymax=emmean+SE), width=0.2, position=pd) +
  coord_cartesian(ylim=c(0,15000)) + # this is way too much white space
  NULL
gg1
```

The plot with unconditional SE (raw) reproduces

## Oxidative stress in offspring

Table 2 cannot be reproduced because two of the necessary columns (Sex and Plate) are not available (that I can see) in the Excel file (the analysis is a linear mixed model with Block and Plate as random factors and Treatment and Sex as fixed factors).

```{r echo=FALSE, eval=FALSE}
m1 <- lme(Liver_Thiols ~ Treatment*Sex, random = ~ 1|Block, method="ML", data=exp5)
anova(m1)
```

```{r, echo=FALSE, eval=FALSE}
fn <- "table2.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

# Statistical Red Flags

## 1. Computation of ANOVA table using Afex package
The afex package replicates the SPSS output for the growth rate analysis but the design is very unbalanced. It's unclear to me how afex handles this; is it like AOV (which doesn't handle unbalanced data well) or more like anova(lme/lmer), which does handle unbalanced data well (with caveats on Type III interpretation). So I compare the SPSS-like repeated measures anova with lme/lmer and also with simple fixed effects analysis (lm and aov)

### How balanced is the design
```{r}
# how balanced is design?
exp1_long[!is.na(Weight), .(N=.N), by=.(Treatment, Sex, Time)]
```

### ANOVA using Afex (computation above)

```{r}
fit.aov_4
```

##3 ANOVA table using lme/lmer
```{r}
# model III anova of lme/lmer fis
contrasts(exp1_long$Treatment) <- contr.sum
contrasts(exp1_long$Sex) <- contr.sum
m1 <- lme(Weight ~ Treatment*Sex*Time, random = ~ 1|ID, data=exp1_long, method = "ML")
anova(m1, type="marginal")
m2 <- lmer(Weight ~ Treatment*Sex*Time + (1|ID), data=exp1_long, REML = FALSE)
anova(m2, type="marginal")
contrasts(exp1_long$Treatment) <- contr.treatment
contrasts(exp1_long$Sex) <- contr.treatment
```

If we're focussing on p-values, then 1) the big difference using lme/lmer compared to Afex (SPSS/SAS like repeated measures anova) is the p-value for Treatment but also 2) another difference is that for Treatment x Sex. That said, it doesn't make much sense to interpret the Treatment p-value as evidence for "an effect" given the interactions.

### Compare LMM to ANOVA table using lm() and AOV (ignoring clustering due to repeated measures)
```{r}
# fixed effects only
contrasts(exp1_long$Treatment) <- contr.sum
contrasts(exp1_long$Sex) <- contr.sum
# using linear model
Anova(lm(Weight ~ Treatment*Sex*Time, data=exp1_long), type="3")
# classical ANOVA
summary(aov(Weight ~ Treatment*Sex*Time, data=exp1_long))
contrasts(exp1_long$Treatment) <- contr.treatment
contrasts(exp1_long$Sex) <- contr.treatment
```

### Plot it to visualize the effects
```{r}
# linear mixed model "version" of repeated measures ANOVA 
fit.lme <- lme(Weight ~ Treatment*Sex*Time, random = ~ Time|ID, data=exp1_long, method = "ML")
coef(summary(fit.lme))

fit.emm <- data.table(summary(emmeans(fit.lme, specs=c("Treatment", "Sex", "Time"))))
setnames(fit.emm, old="emmean", new="Weight")
pd <- position_dodge(0.5)
gg <- ggplot(data=exp1_long, aes(x=Treatment, y=Weight, color=Time)) +
  #geom_point(position=pd) +
  geom_point(data=fit.emm, aes(x=Treatment, y=Weight), position=pd)+
  geom_errorbar(data=fit.emm, aes(x=Treatment, ymin=lower.CL, ymax=upper.CL), position=pd, width=0.2) +
  facet_grid(.~Sex) +
  NULL
gg
```

