---
title: Data from The interaction between predation risk and food ration on behavior
  and morphology of Eurasian perch
author: Jeff Walker
date: '2018-06-17'
slug: data-from-the-interaction-between-predation-risk-and-food-ration-on-behavior-and-morphology-of-eurasian-perch
categories: []
tags:
  - anova
  - ecology
  - experiment
  - factorial
  - nested
  - linear mixed model
subtitle: ''
output:
  blogdown::html_page:
    toc: true
---
# Sources
Svanbäck R, Zha Y, Brönmark C, Johansson F (2017) The interaction between predation risk and food ration on behavior and morphology of Eurasian perch. Ecology and Evolution 7(20): 8567-8577. https://doi.org/10.1002/ece3.3330

Svanback R, Zha Y, Brönmark C, Johansson F (2017) Data from: The interaction between predation risk and food ration on behavior and morphology of Eurasian perch. Dryad Digital Repository. https://doi.org/10.5061/dryad.975m1

# Background

## Reproducibility key
1. Reproducibility: 1 (fails to reproduce). Not really close
2. Reproducibility: 2 (reproduces to a similar value). Differences may be due to slightly different specification of model or to slight differences in data (case may be excluded or missing)
3. Reproducibility: 3 (reproduces to decimal or random error). Differences due to rounding or stochastic part of estimation.
4. Reproducibility: 3- (reproduces to something close to random error). Differences very small and due to either different specification or stochastic or maybe algorithmic differences between packages.

A "+" means that the precise method wasn't specified and reproducibility required exploration.

## Caveats
Failure to reproduce (level 1) or partial reproducibility (level 2 or 3-) very well might be an error in my coding, or my misunderstanding of the author's methods, or my lack of knowledge of statistics generally or the R functions that I use more specifically. Similarly, my **Red Flags** may also represent my own lack of knowledge or best practices in statistics.

# Reproducibility
```{r prep, echo=FALSE, message=FALSE}
library(knitr)
library(readxl)
library(ggplot2)
library(data.table)
library(nlme)
library(lme4)
library(lmerTest)
library(emmeans)

base_path <- "../data" #knitr
#base_path <- "content/data"  # console
folder <- '2018-06-17-data-from-the-interaction-between-predation-risk-and-food-ration-on-behavior-and-morphology-of-eurasian-perch'
```

## Figure 1/Table 1 (Condition and Length)

```{r readFig1}
fn <- 'individual data.csv'
file_path <- paste(base_path, folder, fn, sep='/')
dt.fig1 <- fread(file_path)
```

### Condition index
> We estimated condition of each individual by first performing a regression on all log-transformed lengths and weights and saving the residuals. From this regression, we calculated the weight for an average individual of 90 mm (the average length at the end of the experiment was 90.6 mm). Condition for each fish was then estimated as the calculated average weight plus the residual. The condition index was then used as a measure of condition in all analyses. 

I assume that "performing a regression on all log-transformed lengths and weights and saving the residuals" means to regress weight on length.

```{r fig1Condition}
dt.fig1[, ln_weight:=log(Weight)]
dt.fig1[, ln_length:=log(Length)]
fit <- lm(ln_weight ~ ln_length, na.action="na.exclude", data=dt.fig1)
yhat <- predict(fit, newdata = data.frame(ln_length=log(90)))
dt.fig1[, condition:=yhat + residuals(fit)]
dt.fig1[, condition2:=exp(yhat + residuals(fit))]
dt.fig1[, condition3:=exp(yhat) + exp(residuals(fit)) - 1] # sub exp(0)=1

dt.fig1[, .(mean=exp(mean(condition)), # marginal mean then backtransform
            SE=exp(sd(condition)/sqrt(.N)),
            mean2=mean(condition2), # backtransform then mean
            SE2=sd(condition2/sqrt(.N)),
            mean3=mean(condition3), # backtransform yhat and residuals then mean
            SE3=sd(condition3/sqrt(.N)) 
            ), by=c('Pike_presence', 'Food_Ratio')]
```

mean computes the statistics in the transformed space and then backtransforms, which makes a little more sense. condition3/mean3 is the closest to the values suggest by Figure 1 but the range of my means is slightly larger.

Reproducibility: 2+ (reproduces to a similar value).

### Table 1

it's much simpler for me to think of this as factorial design with a random (aquarium) or clustered component than to think of the tanks as being nested within Pike_presence*Food_Ratio. Tank is the cluster unit.

```{r table1}
# condition
model <- formula(condition3 ~ Pike_presence*Food_Ratio)
fit.condition <- lme(model, random = ~ 1|Aquarium_ID, data=dt.fig1)
anova(fit.condition, type='sequential') # Type I
anova(fit.condition, type='marginal') # Type III
# type II for Food_Ratio
model <- formula(condition3 ~ Pike_presence + Food_Ratio)
anova(lme(model, random = ~ 1|Aquarium_ID, data=dt.fig1), type='sequential')
# type II for Pike_presence
model <- formula(condition3 ~ Food_Ratio + Pike_presence)
anova(lme(model, random = ~ 1|Aquarium_ID, data=dt.fig1), type='sequential')

# length
model <- formula(Length ~ Pike_presence*Food_Ratio)
fit.length <- lme(model, random = ~ 1|Aquarium_ID, data=dt.fig1)
anova(fit.length, type='sequential') # Type I
anova(fit.length, type='marginal') # Type III
# type II for Food_Ratio
model <- formula(Length ~ Pike_presence + Food_Ratio)
anova(lme(model, random = ~ 1|Aquarium_ID, data=dt.fig1), type='sequential')
# type II for Pike_presence
model <- formula(Length ~ Food_Ratio + Pike_presence)
anova(lme(model, random = ~ 1|Aquarium_ID, data=dt.fig1), type='sequential')
```

Condition: Reproducibility: 2+ (reproduces to a similar value). Probably a difference in computation of Condition index. Reproduces uses sequential (default) but not marginal SS.

Length: Reproducibility: 3+ (reproduces to decimal or random error). Reproduces uses sequential (default) but not marginal SS.


**Red Flag**. Table 1 reports Type 1 (sequential) sums of squares and almost certainly the authors want Type III SS given the interaction effect in the model. Notice this makes a key difference here if inference is based on alpha.

### Tukey Contrasts

```{r fig1Contrasts}
inc_rows <- c('y,05_low - n,05_low', 'y,10_medium - n,10_medium', 'y,15_high - n,15_high')
condition.emm <- emmeans(fit.condition, specs=c('Pike_presence', 'Food_Ratio'))
condition.contrasts <- data.table(summary(contrast(condition.emm, method='revpairwise')))
condition.contrasts[contrast %in% inc_rows,]

length.emm <- emmeans(fit.length, specs=c('Pike_presence', 'Food_Ratio'))
length.contrasts <- data.table(summary(contrast(length.emm, method='revpairwise')))
length.contrasts[contrast %in% inc_rows,]

```

Condition post-hoc: Reproducibility: 1+ (fails to reproduce). Compounded difference due to difference in Condition index mean and standard error. My pattern of mean and SE are similar and close in values but do not replicate at level 3. 

Length post-hoc: Reproducibility: NA. Qualitatively it reproduces but numbers aren't given in text.

## Figure 2/Table 2 (Behavior)

```{r readFig2}
fn <- 'behavioral data.csv'
file_path <- paste(base_path, folder, fn, sep='/')
dt.fig2 <- fread(file_path)
```

### Table 2

I assume "distance from the partitioning" is the distance_from_predator in the table.
```{r}
# rename columns to remove spaces
setnames(dt.fig2, old=c('distance from the partitioning', 'distance from the bottom', 'time_measured(start or end of experiment)'),
         new=c('distance_from_predator', 'distance_from_bottom', 'time'))

# start of experiment
# distance from predator
model <- formula(distance_from_predator ~ Pike_presence*Food_Ratio)
fit.condition <- lme(model, random = ~ 1|Aquarium_ID, data=dt.fig2[time=='start'])
anova(fit.condition, type='sequential') # Type I
anova(fit.condition, type='marginal') # Type III

# distance from bottom
model <- formula(distance_from_bottom ~ Pike_presence*Food_Ratio)
fit.condition <- lme(model, random = ~ 1|Aquarium_ID, data=dt.fig2[time=='start'])
anova(fit.condition, type='sequential') # Type I
anova(fit.condition, type='marginal') # Type III

# swimming distance
model <- formula(tot_dist_swimming ~ Pike_presence*Food_Ratio)
fit.condition <- lme(model, random = ~ 1|Aquarium_ID, data=dt.fig2[time=='start'])
anova(fit.condition, type='sequential') # Type I
anova(fit.condition, type='marginal') # Type III

# end of experiment
# distance from predator
model <- formula(distance_from_predator ~ Pike_presence*Food_Ratio)
fit.condition <- lme(model, random = ~ 1|Aquarium_ID, data=dt.fig2[time=='end'])
anova(fit.condition, type='sequential') # Type I
anova(fit.condition, type='marginal') # Type III

# distance from bottom
model <- formula(distance_from_bottom ~ Pike_presence*Food_Ratio)
fit.condition <- lme(model, random = ~ 1|Aquarium_ID, data=dt.fig2[time=='end'])
anova(fit.condition, type='sequential') # Type I
anova(fit.condition, type='marginal') # Type III

# swimming distance
model <- formula(tot_dist_swimming ~ Pike_presence*Food_Ratio)
fit.condition <- lme(model, random = ~ 1|Aquarium_ID, data=dt.fig2[time=='end'])
anova(fit.condition, type='sequential') # Type I
anova(fit.condition, type='marginal') # Type III

```

Reproducibility: 3+ (reproduces to decimal or random error). Reproduces uses sequential (default) but not marginal SS.

**Red Flag**. Table 2 reports Type 1 (sequential) sums of squares and almost certainly the authors want Type III SS (or Type II) given the interaction effect in the model. Notice this makes a key difference here if inference is based on alpha.

## Morphometrics

# Post-publication review

As noted above, the ANOVA tables reported present Type I (sequential) SS which are almost certainly not what the authors intended as the p-values of the first factor will not be conditional on the second factor (nor the interaction). This has consequences on the inference from the ANOVA tables, which raises the question of the role of "storytelling" to "explain" whatever results occur. Would the revised tables generate the same "explanatory" story? It's one reason why I just don't find ANOVA tables worthwhile (at all). I'd much rather use sets of contrasts to understand what is going on and at least these also give me effect sizes.