---
title: Data from Ontogenetic thermal tolerance and performance of ectotherms at variable
  temperatures
author: Jeff Walker
date: '2018-05-26'
slug: data-from-ontogenetic-thermal-tolerance-and-performance-of-ectotherms-at-variable-temperatures
categories: []
tags:
  - experiment
  - 2x2
  - physiology
subtitle: ''
output:
  blogdown::html_page:
    toc: true
---
# Sources

Cavieres G, Bogdanovich JM, Bozinovic F (2016) Ontogenetic thermal tolerance and performance of ectotherms at variable temperatures. Journal of Evolutionary Biology 29(7): 1462-1468. https://doi.org/10.1111/jeb.12886. [Google Scholar](https://scholar.google.com/scholar?hl=en&as_sdt=0%2C20&q=Ontogenetic+thermal+tolerance+and+performance+of+ectotherms+at+variable&btnG=){target="_blank"}

Cavieres G, Bogdanovich JM, Bozinovic F (2016) Data from: Ontogenetic thermal tolerance and performance of ectotherms at variable temperatures. Dryad Digital Repository. [https://doi.org/10.5061/dryad.dm93h]( https://doi.org/10.5061/dryad.dm93h){target="_blank"}

# Background

Explored changes in thermal performance curve in response to rearing at 1 of 4 temperature classes (2 mean temperatures x 2 temperature fluctations). Compared to Huey & Kingsolver's three evolutionary responses, so considered this an ontogenetic response. The three kinds of responses are 1) right shift in upper thermal limit (UTL), 2) right shift in whole curve -- so increase in UTL but decrease in lower thermal limit (LTL) - this is a trade-off, 3) increase in both UTL and LTL, which broadens the thermal performance (Thermal breath (Tbr) and possibly Optimal Range (OR))

Just looking at their Fig. 2, the response was a right shift of the whole curve. They used a model selection procedure to reduce to a minimum adequate model but this is a case where they want estimates for LTL, UTL, TB, and OR to compare to the H&K models!


## Reproducibility key
1. Reproducibility: 1 (fails to reproduce). Not really close
2. Reproducibility: 2 (reproduces to a similar value). Differences may be due to slightly different specification of model or to slight differences in data (case may be excluded or missing)
3. Reproducibility: 3 (reproduces to decimal or random error). Differences due to rounding or stochastic part of estimation.
4. Reproducibility: 3- (reproduces to something close to random error). Differences very small and due to either different specification or stochastic or maybe algorithmic differences between packages.

A "+" means that the precise method wasn't specified and reproducibility required exploration.

## Caveats
Failure to reproduce (level 1) or partial reproducibility (level 2 or 3-) very well might be an error in my coding, or my misunderstanding of the author's methods, or my lack of knowledge of statistics generally or the R functions that I use more specifically. Similarly, my **Red Flags** may also represent my own lack of knowledge or best practices in statistics.

# Summary of Re-analysis
## Reproducibility

Table 1 - mostly reproducible but some rounding error.

Table 2 - not reproducible. The methods aren't clear on the model to compute the marginal means and I cannot recover them.

## Post-publication review
Their Fig. 2 is good but their Tables 1 and 2 are not the tables needed to make the claims they do. Where were the reviewers? I add this. There is one red flag based on anomalous data in Fig. 1 that suggests (contrary to authors) that there is a right shift in $UTL$ in the constant-$T$ group.

# Reproducibility
```{r libraries, echo=TRUE, message=FALSE}
library(knitr)
library(readxl)
library(ggplot2)
library(doBy)
library(data.table)
library(MuMIn)
library(emmeans)
library(cowplot)
library(HarrellPlot)
```

```{r data}
base_path <- "../data" #knitr
#base_path <- "content/data"  # console

folder <- '2018-05-26-data-from-ontogenetic-thermal-tolerance'
fn <- 'data_cavieres et al JEB 2016.xls'
file_path <- paste(base_path,folder, fn, sep='/')
dt <- data.table(read_excel(file_path))
old_colnames <- colnames(dt)
ycols <- c('Tbar', 'Tfluc', 'Sex', 'X1', 'X2', 'OR', 'To', 'Vmax', 'Tbr', 'LTL', 'UTL')
setnames(dt, old_colnames, ycols)
dt[,Tbar:=factor(Tbar, c('15','28'))]
dt[,Tfluc:=factor(Tfluc, c('0','4'))]
# order by factor for HarrelPlot
dt <- orderBy(~ Tbar + Tfluc, dt)
```

The authors used a model selection method to get something like minimum adequate model for the effects of $Sex$ + $Tbar$ + $Tfluc$ on 8 performance curve traits. The model selection method was

>To test the potential effects of mean of temperature and daily fluctuation, we performed a linear model, with following variables: sex + mean of temperature * daily fluctuation. Model selection was performed using Akaike’s information criterion for finites sample (AICc); both delta AIC (Di) values, a measure of each model relative to the best model, and model weights (wi), a measure of the evidence supporting a specific model, were used for selections and model comparison. When there were multiple models with Di < 2, selection was based on the difference in parameters present in each model, reduction in deviance and log-likelihood values (Burnham & Anderson, 2003)

"selection was based on the difference in parameters present in each model, reduction in deviance and log-likelihood" is not reproducible language. I don't know what this means but this is a critical part of the methods? Did the reviewers understand this or did they just give it a pass?

## Table 1

```{r table1}
ycols <- c('To', 'Vmax', 'Tbr', 'LTL', 'UTL', 'OR', 'X1', 'X2')
table1 <- list()
for(j in 1:length(ycols)){
  full_model <- formula(paste(ycols[j], "~ Sex + Tbar*Tfluc"))
  fit <- lm(full_model, data=dt, na.action=na.fail)
  fit.all <- dredge(fit)
  table1[[j]] <- round(coefficients(summary(get.models(fit.all, subset=1)[[1]])), 3)
}
names(table1) <- ycols
table1
```

Table 1 reproducibility

1. $T_o$ reproduces
2. $Vmax$ cannot recover
3. $Tbar$ reproduces
4. $LTL$ reproduces *except* sex effect. The authors seem to have merely not reported the Sex effect -- otherwise how to explain equivalence of coefficients, SE, and T
5. $UTL$ reproduces
6. $OR$ reproduces
7. $X1$ reproduces
8. $X2$ reproduces
 
## Table 2

These seem to be estimated marginal means, presumably from the models in Table 1. But this can't be because 28C-15 != 28V-15V for $T_o$ and others. Also, $OR$ is not simply the intercept (and must have an interaction too).

Tried various combinations of the IVs. An additive model ignoring sex is similar but again, cannot be it because there is some kind of interaction in their results, just not the modeled interaction.

```{r table2}
dredge_it <- FALSE
table2 <- list()
for(j in 1:length(ycols)){
  full_model <- formula(paste(ycols[j], "~ Tbar + Tfluc"))
  fit <- lm(full_model, data=dt, na.action=na.fail)
  if(dredge_it){
    fit.all <- dredge(fit)
    fit <- get.models(fit.all, subset=1)[[1]]
    best_vars <- all.vars(formula(fit))[-1] # if using dredge
    if(length(best_vars > 0)){
      table2[[j]] <- emmeans(fit, specs=best_vars)
    }else{
      table2[[j]] <- coefficients(fit)[1]
    }    
  }else{
      table2[[j]] <- emmeans(fit, specs=c('Tbar', 'Tfluc'))
  }
}
names(table2) <- ycols
table2

```

Table 2 reproducibility -- partial.

Regardless, this makes little difference to the inference.

## Table 3

Is this correlated change across the trial temperatures? There is no data for this!

## Figure 2

The data to re-plot Figure 2 were not made available.

# Post-publication review
## Use the full model to compare to Huey & Kinsolver

To compare to H&K models, we need estimates for change in $LTL$, $UTL$, $Tbr$ and possibly $OR$ across treatment effects. If the interaction is very smallish, one could report the pooled effects across both levels of the other treatment. Here I report the simple effects (differences within each level of the other treatment).

```{r fullmodels}
ycols <- c('To', 'Vmax', 'Tbr', 'LTL', 'UTL', 'OR', 'X1', 'X2')
table_list <- list()
gg_list <- list()
for(j in 1:length(ycols)){
  
  res <- HarrellPlot(x="Tfluc", y=ycols[j], g="Tbar", data=dt, 
                     add_interaction=TRUE, interaction.group = TRUE, 
                     contrasts.method='revpairwise')
  table_list[[j]] <- res$tables$contrasts
  gg_list[[j]] <- res$gg
}
names(table_list) <- ycols
names(gg_list) <- ycols
```

### Table of contrasts needed for interpretation
```{r fullModelTables}
table_list
```

### Harrell Plots of full model ignoring sex
```{r fullModelPlots}
gg_list
```

### Summary of my Harrell Plots and Table of contrasts
1. $T_o$

2. $Tbr$ is noisy relative to signal but the pattern is small decrease in $Tbr$ given temperature flucation but no change in mean and small increase in $Tbr$ given change in temperature regardless of fluctuation.

3. $LTL$ increases with mean temp in both constant and fluctuating T levels but simple fluctuation (with no increase in mean temperature) has effectively zero response.

4. $UTL$ increase with mean temp in fluctuation level but only small increase with mean temp in no-fluctuation level. Increase with fluctuation at both mean temperature levels.

5. $OR$

## Shift in UTL
> in variable environments the upper thermal limit of performance curve shifted to the right, but flies reared in constant environments did not exhibit changes in UTL and the curve did not shift.

Note that their Fig. 2 shows a clear right shift even in the non-fluctuating environment but an anomaly with the data at 40 degrees where the high temperature treatment is unusually low.

Their Table 1 doesn't present the evidence for this -- indeed their presented model does not include the interaction, which suggests a uniform "right shift" (increase in $UTL$) independnet of level of $Tfluc$. But the conclusion implies an interaction. Their Table 2 doesn't present this because it doesn't seem to include the modeled interaction even though there is some small interaction since the contrasts within each level are not equal. Also, a table of marginal means + SE does is not what we want anyway, we want a table of the contrasts (my version of table 2)

The HarrellPlot of the model with interaction shows the result nices - increased $UTL$, with increased $Temperature$, in the fluctuating temperature group but trivially small change in the constant temperature group. The top two contrasts are the $Tbar$ effect, which is the evidence for the shift to right in fluctuating environment vs. non fluctuating environment. The bottom two contrasts are the $Tfluc$ effect. The proper table to show these results are in my version of Table 1.

# Red Flags
1. Fig 1 shows a right shift in UTL in the constant $T$ group but the anomalously low walking speed at 40 C seems to have effected the estimate of $UTL$ (I don't know how this was estimated from the data but the authors give a reference). The fact that this is an anomalously low speed in both groups suggest that it was a systematic error in the measurement (maybe that accidentally increased the to 44 C).

# Statistical misconceptions
1. "Nevertheless, Tbr was not affected by environmental variability." P > 0.05 != "no effect"