---
title: 'Data from: The significance of prey avoidance behaviour for the maintenance
  of a predator colour polymorphism'
author: Jeff Walker
date: '2019-02-02'
slug: data-from-the-significance-of-prey-avoidance-behaviour-for-the-maintenance-of-a-predator-colour-polymorphism
categories: []
tags:
  - generalized linear mixed model
  - binomial response
subtitle: ''
---

# Sources

Ajuria Ibarra H, Kinahan M, Marcetteau J, Mehigan AJR, Ziegelmeier RO, Reader T (2018) The significance of prey avoidance behaviour for the maintenance of a predator colour polymorphism. Behavioral Ecology, online in advance of print. https://doi.org/10.1093/beheco/ary129

Ajuria Ibarra H, Kinahan M, Marcetteau J, Mehigan AJR, Ziegelmeier RO, Reader T (2018) Data from: The significance of prey avoidance behaviour for the maintenance of a predator colour polymorphism. Dryad Digital Repository. https://doi.org/10.5061/dryad.792r3k1

# Summary

# Setup
```{r warning=FALSE, message=FALSE}
library(ggplot2)
library(readxl)
library(data.table)
library(lme4)
library(lmerTest)
library(emmeans)
library(cowplot)
```


```{r}
data_folder <- "../data"
folder <- "Data from The significance of prey avoidance behaviour for the maintenance of a predator colour polymorphism"

# images need to be moved to /static/images/data from.../image.png
knitr_path <- FALSE # run = TRUE to debug

```

# Import
```{r}
fn <- "Data from prey avoidance and predator colour polymorphism.xlsx"
sheet_i <- "Experiment 1"
file_path <- paste(data_folder, folder, fn, sep="/")
exp1 <- data.table(read_excel(file_path, sheet=sheet_i))
exp1[, Flower:=factor(Flower)]
exp1[, Spider:=factor(Spider)]
```


# Reproducibility
## Expeiment 1
### statistical methods
> Treatments were performed in temporal blocks, with each combination offlower species and treatment being applied once in each block. Spiders and treatment-flower combinations were chosen randomly within blocks.

>Honeybee responses in Experiment 1 were analyzed with generalized linear mixed models (GLMMs) with binomial errors and block as a random effect. Binomial responses were: 1) accepted versus not accepted (reject + ignore), 2) accepted versus rejected (with ignored flowers excluded), and 3) inspected (accept + reject) versus not inspected (ignore). Differences between spider treatments were analyzed by comparing a model containing all 4 treatments to a model where the 3 spider treatments were collapsed into one and contrasted with the control.

### cleaning and recoding

1. There is no column identifying block; I will assume that the data are organized into the blocks.

2. The key for the codes in the column "Response" is not given. The responses are "ignored", "reject", "accept" and the codes are 1, 2, 3. It took some trial-and-error to match the code with the response, which must be 1=accept, 2=reject, 3, = ignore

2a. Response[952] has a code value of 13. trial-and-error recoding this to 1 reproduced exactly two of three F values in the first row of Table 1.

3. I attempted to reproduce using three binomial models, one for each response category. This required recoding the "response" column to three separate columns.

4. The column Spider also needed to be recoded to binary (control, not control) for the ANOVA.

```{r recode exp1}
n_combos <- length(levels(exp1$Flower)) * length(levels(exp1$Spider))
n_blocks <- nrow(exp1)/n_combos
exp1[, block:=rep(1:n_blocks, each=n_combos)]

# trial-and-error for item 952
inc <- which(exp1$Response==13)
exp1[inc, Response:=1] 

exp1[, R1:=ifelse(Response==1, 1, 0)] # 1= accept, 0 = not accept
exp1[, R2:=ifelse(Response==1, 1, 
                  ifelse(Response==2, 0, NA))] # 1= accept, 0 = reject, NA = ignore
exp1[, R3:=ifelse(Response==3, 0, 1)] # 1 = inspect (reject or accept), 0 = ignore


exp1[, Spider_binary:=ifelse(Spider=="control", "control", "spider")]
```

```{r trial-and-error, echo=FALSE, eval=FALSE}
exp1[inc, R1:=1] 
m1.R1 <- glmer(R1 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m1.R1, test="Chisq")
m2.R1 <- glmer(R1 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m2.R1, test="Chisq")

exp1[inc, R1:=0] 
m1.R1 <- glmer(R1 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m1.R1, test="Chisq")
m2.R1 <- glmer(R1 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m2.R1, test="Chisq")

exp1[inc, R1:=NA] 
m1.R1 <- glmer(R1 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m1.R1, test="Chisq")
m2.R1 <- glmer(R1 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
drop1(m2.R1, test="Chisq")

```

### Table 1

```{r}

m1.R1 <- glmer(R1 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
(test1a <- drop1(m1.R1, test="Chisq"))
m2.R1 <- glmer(R1 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
(test1b <- drop1(m2.R1, test="Chisq"))

m1.R2 <- glmer(R2 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
(test2a <- drop1(m1.R2, test="Chisq"))
m2.R2 <- glmer(R2 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
(test2b <- drop1(m2.R2, test="Chisq"))

m1.R3 <- glmer(R3 ~ Flower*Spider_binary + (1|block), family=binomial, data=exp1)
(test3a <- drop1(m1.R3, test="Chisq"))
m2.R3 <- glmer(R3 ~ Flower + Spider_binary + (1|block), family=binomial, data=exp1)
(test3b <- drop1(m2.R3, test="Chisq"))

```

### reported table 1
```{r table1, echo=FALSE}
fn <- "table1.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

#### my table 1

```{r mytable1, echo=FALSE}
spider <- c(test1b["Spider_binary", "LRT"], test2b["Spider_binary", "LRT"], test3b["Spider_binary", "LRT"])
flower <- c(test1b["Flower", "LRT"], test2b["Flower", "LRT"], test3b["Flower", "LRT"])
inter <- c(test1a["Flower:Spider_binary", "LRT"], test2a["Flower:Spider_binary", "LRT"], test3a["Flower:Spider_binary", "LRT"])
responses <- c("Accepted v not accepted","Accepted v rejected", "Inspected v not inspected")
table1 <- data.table(Response=responses,
                     Spider=spider,
                     Flower=flower,
                     Interaction=inter)
knitr::kable(table1, digits=c(NA, 3, 3, 3))
```

Effectively reproduces. The first two numbers in row one are exact. Differences could be to different algorthms between versions. I don't think they are due to differences in coding.

### Effect of spider treatments

>No significant differences were found among spider treatments: the effect of collapsing the spider treatments together in the statistical model was not significant for any of the response variables: accepted versus not accepted (χ2 = 2.718, df = 6, P = 0.843), accepted versus rejected (χ2 = 3.477, df = 6, P = 0.747), and inspected versus not inspected (χ2 = 6.484, df = 6, P = 0.371).

```{r}
m3.R1 <- glmer(R1 ~ Flower*Spider + (1|block), family=binomial, data=exp1)
anova(m1.R1, m3.R1, test="Chisq") # this reproduces

m3.R2 <- glmer(R2 ~ Flower*Spider + (1|block), family=binomial, data=exp1)
anova(m1.R2, m3.R2, test="Chisq")

m3.R3 <- glmer(R3 ~ Flower*Spider + (1|block), family=binomial, data=exp1)
anova(m1.R3, m3.R3, test="Chisq")
```

"accepted versus not accepted" reproduces exactly; the other two are close

### Figure 1

```{r}
m3_R1.emm <- emmeans(m3.R1, specs=c("Flower", "Spider"), type="response")
m3_R2.emm <- emmeans(m3.R2, specs=c("Flower", "Spider"), type="response")
m3_R3.emm <- emmeans(m3.R3, specs=c("Flower", "Spider"), type="response")

pd <- position_dodge(0.9)
gg1 <- ggplot(data=summary(m3_R1.emm), aes(x=Flower, y=prob, fill=Spider)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax=asymp.UCL), position=pd, width=0.2) +
  coord_cartesian(ylim=c(0, 0.8)) +
  ggtitle("Accept") +
  NULL
gg2 <- ggplot(data=summary(m3_R2.emm), aes(x=Flower, y=prob, fill=Spider)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax=asymp.UCL), position=pd, width=0.2) +
  coord_cartesian(ylim=c(0, 0.8)) +
  ggtitle("Accept after inspection") +
  NULL
gg3 <- ggplot(data=summary(m3_R3.emm), aes(x=Flower, y=prob, fill=Spider)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax=asymp.UCL), position=pd, width=0.2) +
  coord_cartesian(ylim=c(0, 0.8)) +
  ggtitle("Inspect") +
  NULL

# See myfig1 chunk for plot
```

```{r fig1, echo=FALSE}
fn <- "fig1.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

```{r myfig1, fig.asp = 1.5, echo=FALSE}
plot_grid(gg1, gg2, gg3, nrow=3)
```

The CIs differ. My CIs are computed from the model fit using emmeans. These are much more uniform than the reported CIs. The reported CIs may be from confint(fit_model)

# Statistical Red Flags