---
title: 'Data from: Perceived threats of infanticide reduce maternal allocation during
  lactation and lead to elevated oxidative damage in offspring'
author: Jeff Walker
date: '2019-02-18'
slug: data-from-perceived-threats-of-infanticide-reduce-maternal-allocation-during-lactation-and-lead-to-elevated-oxidative-damage-in-offspring
categories: []
tags:
  - linear mixed model
  - SPSS
subtitle: ''
output:
  blogdown::html_page:
    toc: true
---

# Sources

Gale T, Garratt M, Brooks RC (2018) Perceived threats of infanticide reduce maternal allocation during lactation and lead to elevated oxidative damage in offspring. Functional Ecology 32(9): 2158-2169. https://doi.org/10.1111/1365-2435.13146

Gale T, Garratt M, Brooks RC (2018) Data from: Perceived threats of infanticide reduce maternal allocation during lactation and lead to elevated oxidative damage in offspring. Dryad Digital Repository. https://doi.org/10.5061/dryad.762k24f

# Summary

# Setup
```{r libraries, message=FALSE}
library(ggplot2)
library(readxl)
library(data.table)
library(emmeans)
library(nlme)
library(lme4)
library(lmerTest)
library(afex)
```

```{r}
data_folder <- "../data"
folder <- "Data from Perceived threats of infanticide reduce maternal allocation during lactation and lead to elevated oxidative damage in offspring"

# images need to be moved to /static/images/data from.../image.png
knitr_path <- FALSE # run = TRUE to debug

source("../R/clean_label.R") # bookdown
```

# Import
The single sheet contains all experiments in blocks glued wide
```{r}
fn <- "LBE paper data.xlsx"
file_path <- paste(data_folder, folder, fn, sep="/")

# exp1
exp1 <- data.table(read_excel(file_path, range="A1:E185"))
colnames(exp1) <- clean_label(colnames(exp1))
exp1[, Treatment:=factor(Treatment)]

exp2 <- data.table(read_excel(file_path, range="F1:H45"))
colnames(exp2) <- clean_label(colnames(exp2))
exp2[, Treatment:=factor(Treatment)]

```

# Reproducibility
## Offspring weight

>The mean birth weight of novel male treatment pups (1.619± 0.432g) was not significantly different from that of the familiar male treatment (1.628 ± 0.433g; ANOVA: weight: F1, 182= 0.022, p= 0.882; litter size: F1,42= 0.070, p= 0.792).

```{r}
# birth weight
exp1[,.(birth=mean(birth),
        SE=sd(birth)/sqrt(.N)), by=Treatment]
anova(lm(birth ~ Treatment, data=exp1))

# litter size
anova(lm(Littersize ~ Treatment, data=exp2))

```
Reproduces

and identifies LBE = novel male treatment, MC = familiar male treatment

>To analyse the differences in offspring growth rate we used repeated measures ANOVA reporting within-subjects effects. Mauchly's Test of Sphericity indicated that the assumption of sphericity had been violated (χ2(2) = 34.210, p = <0.001), therefore degrees of freedom were corrected using Greenhouse-Geisser adjusted degrees of freedom. There was a difference in growth rate between the treatments but there was no effect of sex (treatment: F1.7,310=7.49, p= 0.001; interaction between treatment and time: F1.7,310=5668, p= <0.001; sex x treatment: F1.6,306=1.3, p= 0.270).

```{r}
exp1[, ID:=1:nrow(exp1)]
exp1_long <- melt(exp1, id.vars = c("Treatment", "Sex", "ID"), variable.name="Time", value.name = "Weight")
exp1_long[, ID:=factor(ID)]
(fit.car <- aov_car(Weight ~ Treatment*Sex*Time + Error(ID/Time), data=exp1_long))
#(fit.ez <- aov_ez("ID", "Weight", exp1_long, between=c("Treatment", "Sex"), within="Time"))
# (fit_nice <- aov_ez("id","value",obk.long,between=c("treatment"),within=c("phase")))
#summary(fit.ez)
```

# Statistical Red Flags