---
title: 'Data from: Interactions among predators and plant specificity protects herbivores
  from top predators'
author: Jeff Walker
date: '2019-01-13'
slug: data-from-interactions-among-predators-and-plant-specificity-protects-herbivores-from-top-predators
categories: []
tags: []
subtitle: ''
output:
  blogdown::html_page:
    toc: true
---

Source: Bosc C, Roets F, Hui C, Pauw A (2018) Interactions among predators and plant specificity protect herbivores from top predators. Ecology 99(7): 1602-1609. https://doi.org/10.1002/ecy.2377
Additionally, please cite the Dryad data package:

Data source: Bosc C, Roets F, Hui C, Pauw A (2018) Data from: Interactions among predators and plant specificity protects herbivores from top predators. Dryad Digital Repository. https://doi.org/10.5061/dryad.6k939m4

# Summary

1. Very clean data files, readme files, and R script.
2. My script is independent and was written prior to opening the author's script (they are in fact quite different, which is typical of R scripts)
3. Only analyzed the experiment results and not the PCA results
4. Estimated marginal means and SEs do NOT reproduce for two reasons, 1) the methods state that *Slope* is in the model but the reported results are from a model not including *Slope*, and 2) the standard errors of the mean for the non-reference levels (Open and Shade) are the standard errors of the difference of these levels from the reference level (Cage) and not the standard errors of the means.
5. The ANOVA table results reproduce exactly.
6. Model checking suggests the data should be analyzed by generalized LMM *if* the goal is parameter estimation.

# Set up
```{r setup, warning=FALSE, message=FALSE}
library(stringr)
library(ggplot2)
library(ggpubr)
library(lmerTest)
library(emmeans)
library(data.table)

data_folder <- "../data"
folder <- "Data from Interactions among predators and plant specificity protects herbivores from top predators"

# images need to be moved to /static/images/data from.../image.png
knitr_path <- FALSE
```

# Import

Note, the unit identifier is not the same string but has the form "6-C-7TR" in `exp_1` and "X6.C.7TR" in `abundance`

```{r import}
# design
fn <- "ExpeDesign_BirdExpe_Dec2014.csv"
file_path <- paste(data_folder, folder, fn, sep="/")
exp_1 <- fread(file_path, header=TRUE)
setnames(exp_1, old="V1", new="id")

# arthropod abundance
fn <- "ArthropodAbundances_BirdExpe_Dec2014.csv"
file_path <- paste(data_folder, folder, fn, sep="/")
abundance <- fread(file_path, header=TRUE)

#create id column of row id with same format as that in exp_1
abundance[, id:=V1]
abundance[, id:=str_replace(V1,"X","")]
abundance[, id:=gsub('.', '-', id, fixed=TRUE)]
# abundance$id

# file for trophic identification for each arthropod
fn <- "ArthropodTrophicGroup_BirdExpe_Dec2014.csv"
file_path <- paste(data_folder, folder, fn, sep="/")
trophic <- fread(file_path, header=TRUE)
setnames(trophic, old="V1", new="taxon")

# file for taxonomic identification for each arthropod
fn <- "ArthropodTaxonomy_BirdExpe_Dec2014.csv"
file_path <- paste(data_folder, folder, fn, sep="/")
taxonomic <- fread(file_path, header=TRUE)
setnames(taxonomic, old="V1", new="taxon")

```

Need to compute abundances within trophic levels and taxonomic levels and then merge with the treatment levels in `exp_1`. 

```{r merge counts into exp_1}
taxa <- setdiff(colnames(abundance), c("V1", "id"))
# sum rows (count arthropods in each site)
abundance[, total_n:=apply(abundance[, .SD, .SDcols=taxa], 1, sum)]

predator_cols <- trophic[TrophicGrp=="Pred", taxon]
abundance[, pred_n:=apply(abundance[, .SD, .SDcols=predator_cols], 1, sum)]

herb_cols <- trophic[TrophicGrp=="Herb", taxon]
abundance[, herb_n:=apply(abundance[, .SD, .SDcols=herb_cols], 1, sum)]

# merge abundance into exp_1
y_cols <- c("id", "total_n", "pred_n", "herb_n")
exp_1 <- merge(exp_1, abundance[, .SD, .SDcols=y_cols], by="id")

# save this for adding columns and debugging
working_exp_1 <- copy(exp_1)
```

# Reproducibility

## Effect on total predators

```{r total-predator}
# "Significance of the fixed factors were determined with REML t-tests and Satterthwaite's approximation on the intercept in each model (package lmerTest in R software)"

fit.pred <- lmer(pred_n ~ Slope + Treatment + (1|Site), data=exp_1)
pred.anova <- anova(fit.pred) #check Table S1 Appendix S3
my.anova.table <- data.table(response="total predator", pred.anova["Treatment",])
```

### my results
```{r total-predator-my-results}
(pred.emm <- emmeans(fit.pred, specs="Treatment", lmer.df = "satterthwaite"))
```

### reported results
The authors report "intermediate predators were significantly more abundant in the exclosure treatments (estimated mean number of individuals ± SE: 27.9 ± 5.2) compared to both open (14.2 ± 2.6) and shade (16.4 ± 3.2)"

The means are the same but the reported standard errors differ. The reported SEs come from the table of coefficients of the model fit *without* slope (contra methods). 

```{r total-predator-coefficients}
# using coefficients
coef(summary(lmer(pred_n ~ Treatment + (1|Site), data=exp_1)))
```

The reported standard errors of the mean for "open" and "shade" are the the SE of the *differences between means* of each from the exclosure (these are standard errors of differences).

Checking the authors supplied code, they did fit the model without Slope as a fixed effect so I've refit here with the correct standard errors of the means for "Open" and "Shade".

```{r total-predator-emm-corrected}
fit.pred <- lmer(pred_n ~ Treatment + (1|Site), data=exp_1)
(pred.emm <- emmeans(fit.pred, specs="Treatment", lmer.df = "satterthwaite"))
```

and the table of contrasts is, which contrasts are used to generate the asterisks for Fig 1.

```{r total-predator-contrasts}
(pred.contr <- contrast(pred.emm, method="revpairwise"))
```

## Effect on total herbaceous
```{r total-herbaceous}
fit <- lmer(herb_n ~ Slope + Treatment + (1|Site), data=exp_1)
herb.anova <- anova(fit) #check Table S1 Appendix S3
my.anova.table <- rbind(my.anova.table,
                        data.table(response="total herbaceous", herb.anova["Treatment",]))

```

### my results
```{r total-herbaceous-my-results}
(herb.emm <- emmeans(fit, specs="Treatment", lmer.df = "satterthwaite"))
```

### reported results
The authors report: "exclosure: 89 ± 19.1, open: 84.9 ± 13.1, shade: 85.6 ± 16"

Again, the means are the same but the reported standard errors differ. The differences are due to the same two errors in the authors analysis: 1) Slope not included in model per methods and 2) SEDs of Open and Shade used instead of SEMs. Compare:

```{r}
coef(summary(lmer(herb_n ~ Treatment + (1|Site), data=exp_1)))
```


Here is the table of corrected standard error of means for the model fit without Slope.

```{r}
fit <- lmer(herb_n ~ Treatment + (1|Site), data=exp_1)
(herb.emm <- emmeans(fit, specs="Treatment", lmer.df = "satterthwaite"))
```

And the table of contrasts, which is used to get the asterisks for Figure 1

```{r}
(herb.contr <- contrast(herb.emm, method="revpairwise"))
```

## ANOVA table in Appendix 3, Table S1

```{r my-anova}
knitr::kable(my.anova.table, digits=c(NA, 3, 3, 0, 4, 3, 3))
```


```{r tableS1, echo=FALSE}
fn <- "tableS1.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

These reproduce exactly

## Fig 1

```{r fig1, echo=FALSE}
fn <- "fig1.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

## Subsetting the response

Will assume going forward that Slope is *not* in the model. Compare the means and SEs to the plots in Appendix 4, Fig. S6
```{r subsets}
exp_1 <- copy(working_exp_1)
spider_cols <- taxonomic[Order=="Araneae", taxon]
pred_spider_cols <- intersect(predator_cols, spider_cols)
abundance[, spider_n:=apply(abundance[, .SD, .SDcols=pred_spider_cols], 1, sum)]

insect_cols <- taxonomic[Class=="Insecta", taxon]
pred_insect_cols <- intersect(predator_cols, insect_cols)
abundance[, pred_insect_n:=apply(abundance[, .SD, .SDcols=pred_insect_cols], 1, sum)]

Hemiptera_cols <- intersect(herb_cols, taxonomic[Order=="Hemiptera", taxon])
abundance[, hemiptera_n:=apply(abundance[, .SD, .SDcols=Hemiptera_cols], 1, sum)]

Coleoptera_cols <- intersect(herb_cols, taxonomic[Order=="Coleoptera", taxon])
abundance[, coleoptera_n:=apply(abundance[, .SD, .SDcols=Coleoptera_cols], 1, sum)]

Lepidoptera_cols <- intersect(herb_cols, taxonomic[Order=="Lepidoptera", taxon])
abundance[, lepidoptera_n:=apply(abundance[, .SD, .SDcols=Lepidoptera_cols], 1, sum)]

Orthoptera_cols <- intersect(herb_cols, taxonomic[Order=="Orthoptera", taxon])
abundance[, orthoptera_n:=apply(abundance[, .SD, .SDcols=Orthoptera_cols], 1, sum)]


# merge abundance into exp_1
y_cols <- c("id", "spider_n", "pred_insect_n", "hemiptera_n", "coleoptera_n", "lepidoptera_n", "orthoptera_n")
exp_1 <- merge(exp_1, abundance[, .SD, .SDcols=y_cols], by="id")

```

### spiders
```{r spiders}
fit <- lmer(spider_n ~ Treatment + (1|Site), data=exp_1)
emmeans(fit, specs="Treatment")
# compare SEMs from coef table
coef(summary(fit))
```

### predatory insects
```{r predatory insects}
fit <- lmer(pred_insect_n ~ Treatment + (1|Site), data=exp_1)
emmeans(fit, specs="Treatment")
# compare SEMs from coef table
coef(summary(fit))
```

### Hemiptera
```{r hemiptera}
fit <- lmer(hemiptera_n ~ Treatment + (1|Site), data=exp_1)
emmeans(fit, specs="Treatment")
# compare SEMs from coef table
coef(summary(fit))
```

### Coleoptera
```{r coleoptera}
fit <- lmer(coleoptera_n ~ Treatment + (1|Site), data=exp_1)
emmeans(fit, specs="Treatment")
# compare SEMs from coef table
coef(summary(fit))
```

### Lepidoptera
```{r lepidoptera}
fit <- lmer(lepidoptera_n ~ Treatment + (1|Site), data=exp_1)
emmeans(fit, specs="Treatment")
# compare SEMs from coef table
coef(summary(fit))
```

### Orthoptera

```{r orthoptera}
fit <- lmer(orthoptera_n ~ Treatment + (1|Site), data=exp_1)
emmeans(fit, specs="Treatment")
# compare SEMs from coef table
coef(summary(fit))

```

## Fig S6
```{r figS6}
fn <- "figS6.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

# Statistical red-flags

A quick model check of the residuals shows classic count response

1. The normal Q-Q shows 1) smaller than expected quantiles at low end (piled up closer to mean) and 2) way too big quantiles at the big end (right skewed)
2. The sqrt(residuals) vs. fitted plot shows a strong positive increase indicating strong heteroskedasticity (increase in variance with increased conditional mean)

This suggests that if estimates are the goal then the data should be fit with a generalized LMM. But if hypothesis testing is the goal then the normal assumption in the LMM is probably okay.

```{r model check}
pred.residuals <- residuals(fit.pred)
exp_1[, pred.residuals := residuals(fit.pred)]
exp_1[, pred.fitted := fitted(fit.pred)]
ggqqplot(data=exp_1, x = "pred.residuals", title="Normal Q-Q")
qplot(x=pred.fitted, y=sqrt(abs(pred.residuals)), data=exp_1) + geom_smooth(method="lm")
```

