---
title: Data from Organic matter and nutrient inputs from large wildlife influence
  ecosystem function in the Mara River, Africa
author: Jeff Walker
date: '2019-01-29'
slug: data-from-organic-matter-and-nutrient-inputs-from-large-wildlife-influence-ecosystem-function-in-the-mara-river-africa
categories: []
tags:
  - emtrends
subtitle: ''
output:
  blogdown::html_page:
    toc: true
---

# Sources

Subalusky AL, Dutton CL, Njoroge L, Rosi EJ, Post DM (2018) Organic matter and nutrient inputs from large wildlife influence ecosystem function in the Mara River, Africa. Ecology 99(11): 2558-2574. https://doi.org/10.1002/ecy.2509

Subalusky AL, Dutton CL, Njoroge L, Rosi EJ, Post DM (2018) Data from: Organic matter and nutrient inputs from large wildlife influence ecosystem function in the Mara River, Africa. Dryad Digital Repository. https://doi.org/10.5061/dryad.g886d9v

# Summary

1. The files are nearly perfectly clean.

2. Nutrient dissolution: The authors do not use repeated measures anova as specified in the methods. The fixed effects ANOVA reproduce the ANOVA table. The post-hoc t-tests are reproduced qualitatively but not quantitatively.

3. Decomposition rates: The ANOVA statistics reproduce using the column "ln_hippo_prop_rem", which is not the column I inferred from either the methods or from the text description of the results.

4. Nutrient limitation: Reproduces as specified.

5. Ecosystem function: Experimental streams: The response log(GPP) (as specified in methods) does not reproduce but untransformed GPP does reproduce. The response Sqrt(chl) (as specified in methods) does not reproduce but log(chl) does reproduce. The rates of increase reproduce. River biofilms: All analyses reproduce at least qualitatively (I cannot check quantiative results where none were given). Treatment column had to be re-coded in order to reproduce.

6. A few minor to moderate statistical red flags are noted.

# Set up

```{r libraries, message=FALSE, warning=FALSE}
library(broom)
library(ggplot2)
library(ggpubr)
library(data.table)
library(readxl)
library(nlme)
library(emmeans)
```

```{r import path}
data_folder <- "../data"
folder <- "Data from Organic matter and nutrient inputs from large wildlife influence ecosystem function in the Mara River, Africa"

# images need to be moved to /static/images/data from.../image.png
knitr_path <- FALSE # run = TRUE to debug
```

```{r clean labels}
source("../R/clean_label.R") # bookdown

```

# Import
```{r import}
# dissolution rates
fn <- "Subalusky_2018_mara_nutrient_dissolution_chambers.csv"
file_path <- paste(data_folder, folder, fn, sep="/")
dissolution <- fread(file_path)
dissolution[, treatment:=factor(treatment, c("H 1", "H 5", "H 20", "W", "HW"))]

# decomposition rates
fn <- "Subalusky_2018_mara_hippo_feces_decomp_litterbags.csv"
file_path <- paste(data_folder, folder, fn, sep="/")
decomp <- fread(file_path)
colnames(decomp) <- clean_label(colnames(decomp))
# treatment	=	experimental treatment, where A: 40 g fresh hippo feces; B: 10 g aged hippo feces; C: 10 g aged hippo feces + 20 g wildebeest tissue
decomp[, treatment_recode:=ifelse(treatment=="A", "fresh",
                                  ifelse(treatment=="B", "aged", "wb"))]
decomp[, treatment_recode:=factor(treatment_recode, c("fresh", "aged", "wb"))]

# nutrient limitation
fn <- "Subalusky_2018_mara_nutrient_limitation_assays.csv"
file_path <- paste(data_folder, folder, fn, sep="/")
limit <- fread(file_path)

# Ecosystem function
fn <- "Subalusky_2018_mara_experimental_stream_tiles.csv"
file_path <- paste(data_folder, folder, fn, sep="/")
stream <- fread(file_path)

# River biofilms
# note there is another file with the raw data. This is the file of means from that file.
fn <- "Subalusky_2018_mara_river_biofilm_summary_data.csv"
file_path <- paste(data_folder, folder, fn, sep="/")
river <- fread(file_path)

```

# Reproducibility
## Nutrient dissolution rates
### Statistical methods
>We analyzed each of the four response variables (NH4, SRP, DOC and DO) in response to treatment and time using a repeated measures ANOVA with the aov function in R (R Core Team 2018).

Note there is no column for the random factor, which should be chamber, since there were three time measures within each chamber. I created one assuming that the order is uniform.

```{r}
# add chamber ID, this assumes chamber order is maintained
dissolution[, chamber:=paste(treatment, rep(1:3), sep="-")]
```

### ANOVA results
```{r}
summary(aov(log(NH4.no0) ~ treatment*hours + Error(chamber), data=dissolution))$"Error: chamber"
summary(aov(log(SRP.no0) ~ treatment*hours + Error(chamber), data=dissolution))$"Error: chamber"
summary(aov(log(DOC.no0) ~ treatment*hours + Error(chamber), data=dissolution))$"Error: chamber"
```

>There were significant treatment effects of input type and concentration on NH4 (rm ANOVA, F=23.53$_{4,35}$, P<0.001), SRP (F=123.61$_{4,35}$, P<0.001), and DOC (F=27.96$_{4,35}$, P<0.001), and significant treatment by time interaction effects on NH4 (rm ANOVA, interaction, F=9.304,35, P<0.001) and SRP (F=7.594,35, P<0.001).

>There were also significant treatment effects on DO (rm ANOVA, F=101.69$_{4,35}$, P<0.001)

This does not reproduce.

After some exploration, a simple fixed effects model reproduces, so this is not a repeated measures ANOVA, which raises Statistical Red Flag #1 (see below)

```{r}
nh4.fit <- lm(log(NH4.no0) ~ treatment*hours, data=dissolution)
srp.fit <- lm(log(SRP.no0) ~ treatment*hours, data=dissolution)
doc.fit <- lm(log(DOC.no0) ~ treatment*hours, data=dissolution)
do.fit <- lm(log(DO) ~ treatment*hours, data=dissolution)

summary(aov(nh4.fit))
summary(aov(srp.fit))
summary(aov(doc.fit))
summary(aov(do.fit))

```

### Posthoc pairwise comparisons

>If there was a significant treatment effect at α=0.05, we then performed a Tukey pairwise comparison test between treatments using the Holm method to adjust p values for multiple comparisons (Holm 1979, R Core Team 2018).

Note that "Tukey pairwise comparison" is a specific kind of simultaneous posthoc test not the general name for post hoc tests. And, it's not clear what this analysis is, especially given the large interaction for NH4, which makes pairwise comparisons conditional on the level of time. So are the comparisons at 24 hours? Or, was there complete pooling across time?

Here is the Holm adjustment using the pooled sigma from the fixed effect model for NH4. This isn't close to the reported pattern of significance levels.

```{r}
nh4.emm <- emmeans(nh4.fit, specs="treatment")
contrast(nh4.emm, method="revpairwise", adjust="holm") # caution! ignores interaction
```

Here I use pairwise t-test using both the pooled SD over all groups and the pooled SD for the two groups compared. Regardless, this raises Statistical Red Flag #2.

```{r}
pairwise.t.test(x=log(dissolution$NH4.no0), g=dissolution$treatment, p.adjust.method="holm", pool.sd=TRUE)
pairwise.t.test(x=log(dissolution$SRP.no0), g=dissolution$treatment, p.adjust.method="holm", pool.sd=TRUE)
pairwise.t.test(x=log(dissolution$DOC.no0), g=dissolution$treatment, p.adjust.method="holm", pool.sd=TRUE)
```

```{r}
pairwise.t.test(x=log(dissolution$NH4.no0), g=dissolution$treatment, p.adjust.method="holm", pool.sd=FALSE)
pairwise.t.test(x=log(dissolution$SRP.no0), g=dissolution$treatment, p.adjust.method="holm", pool.sd=FALSE)
pairwise.t.test(x=log(dissolution$DOC.no0), g=dissolution$treatment, p.adjust.method="holm", pool.sd=FALSE)

```

note -- the design is balanced so Welch t-test would be the same

> Concentrations of NH4, SRP and DOC increased with increasing hippo feces concentration, although treatment differences were only significant for SRP and DOC (except for low and medium concentrations for DOC) (pairwise t test, P<0.05 for all comparisons).

the qualitative pattern reproduces except: 

using pool.sd=TRUE, DOC:H5 v H20 > 0.05, not H1 v H5

using pool.sd=FALSE, DOC: H5 v H20 < 0.05 and H1 v H5 < 0.05

>The wildebeest treatment was significantly higher in NH4 than the lowest concentration of hippo feces (P=0.017), and it was significantly higher in SRP and DOC than all concentrations of hippo feces (P<0.05 for all comparisons).

qualitatively reproduces except:

using pool.sd=TRUE, NH4: w v H1, p= 0.0073, reported = (P=0.017)

using pool.sd=FALSE, NH4: w v H1, p=0.0023, reported = (P=0.017)

>There were no significant differences between the wildebeest and hippo + wb treatments for NH4 or DOC; however, SRP was higher in the hippo + wb treatment as compared to the wildebeest treatment (p<0.03)

The w vs. wb treatment does not reproduce for SRP
> There was no significant difference in oxygen consumption between the high concentration of hippo feces, wildebeest and hippo + wb treatments.

```{r}
pairwise.t.test(x=log(dissolution$DO), g=dissolution$treatment, p.adjust.method="holm")
```

This pattern reproduces using the non-pooled sigma (independent t-tests).

## Decomposition rates

### Statistical methods
> We used a two-way ANOVA to analyze the influence of treatment (fresh, aged, with wildebeest) on decay rate. Presence of an interaction effect between treatment and time at α=0.05 indicated a significant effect of treatment on decay rate. We then conducted pairwise analyses of each treatment combination using one-way ANOVAs to determine how they differed from one another. We accounted for the number of pairwise comparison tests by using a Bonferroni adjusted α value of 0.0167 to indicate statistical significance. All statistical analyses in this paper were done in R version 3.3.2 (R Core Team 2018).

### ANOVA results
> There was a significant interaction between hippo feces treatment and time on the mass remaining (two-way ANOVA, interaction, F=3.19$_2{},75}$, P=0.047), indicating an effect of treatment on decay rate (Table 1, Appendix S2: Fig. S2). 

```{r}
hippo_post <- lm(hippo_post ~ treatment_recode*days, data=decomp)
summary(aov(hippo_post))
```

This doesn't reproduce. What does reproduce is using the column "ln_hippo_prop_rem", which is the log of the column "adj_hippo_prop_rem". So the response it is not the "mass remaining" but the log of the adjusted proportion of the mass remaining. This raises Statistical Red Flag #3. 

```{r}
decomp[, my_ln_hippo_prop_rem:=log(adj_hippo_prop_rem)]
head(decomp[, .SD, .SDcols=c("ln_hippo_prop_rem", "my_ln_hippo_prop_rem")])

ln_hippo_prop_rem.fit <- lm(ln_hippo_prop_rem ~ treatment_recode*days, data=decomp)
summary(aov(ln_hippo_prop_rem.fit))
```


>Fresh feces decayed faster than aged hippo feces (ANOVA, interaction, F=4.21$_{1,56}$, P=0.045), although this was not significant after accounting for multiple comparisons. There was no significant difference between decay rate of fresh feces and aged feces with carcass (ANOVA, interaction, F=0.18$_{1,49}$, P=0.671); however, aged feces with carcass decayed significantly faster than aged feces alone (ANOVA, interaction, F=8.20$_{1,45}$, P=0.006).

This reproduces (using "ln_hippo_prop_rem") but the independent ANOVAs raises Statistical Red Flag #4

```{r}
fit <- lm(ln_hippo_prop_rem ~ treatment_recode*days, data=decomp[treatment_recode!="wb"])
summary(aov(fit))

fit <- lm(ln_hippo_prop_rem ~ treatment_recode*days, data=decomp[treatment_recode!="aged"])
summary(aov(fit))

fit <- lm(ln_hippo_prop_rem ~ treatment_recode*days, data=decomp[treatment_recode!="fresh"])
summary(aov(fit))
```

## Nutrient limitation
### Statistical method
> Data were square-root transformed to meet assumptions of normality. We used a two-way ANOVA to analyze the response of biofilms to different nutrient treatments at different sites in both 2011 and 2013. If there were significant interactions at α=0.05, we then used one-way ANOVAs to analyze treatment responses separately for each site, followed by Tukey’s HSD post-hoc test for multiple comparisons (R Core Team 2018).

### Respiration Results

> Respiration (R) by heterotrophic biofilms responded differently to nutrient treatments at sites with and without large wildlife inputs in both 2011 (two-way ANOVA, interaction, F=5.40$_{10,68}$, P<0.001) and 2013 (F=3.67$_{12,75}$, P<0.001). 

Reproduces

```{r}
# heterotrophic: substrate = "Sponge"
R_hetero_Y11.fit <- lm(sqrt(CR.corr.no0) ~ treatment*site, data=limit[year=="2011" & substrate=="Sponge"])
summary(aov(R_hetero_Y11.fit))

R_hetero_Y13.fit <- lm(sqrt(CR.corr.no0) ~ treatment*site, data=limit[year=="2013" & substrate=="Sponge"])
summary(aov(R_hetero_Y13.fit))

```

#### Site 1

>At the No Wildlife site in 2011, R on the organic substrate increased significantly relative to the control in response to NH4+PO4 and to NO3+PO4 (one-way ANOVA, F=18.52$_{5,24}, P<0.001; Tukey’s HSD, P=0.005 and P<0.001, respectively), but not in response to any of the nutrients alone, indicating N and P co-limitation (Appendix S2: Fig. S3). 

Reproduces

```{r}
# site key: Site 1="no wildlife", Site 2="Hippo", Site 3="Hippo + wildebeast"
# nutrients added to agar cup; Ctl: control, C: carbon (only measured in 2013), H: ammonium, N: nitrate, P: phosphate, HP: ammonium + phosphate, NP: nitrate + phosphate
fit <- lm(sqrt(CR.corr.no0) ~ treatment, data=limit[year=="2011" & substrate=="Sponge" & site=="Site 1"])
summary(aov(fit))
TukeyHSD(aov(fit))$treatment[c("HP-Ctl", "NP-Ctl"),]
TukeyHSD(aov(fit))$treatment[c("H-Ctl", "N-Ctl", "P-Ctl"),]

```

>At the No Wildlife site in 2013, R increased significantly relative to the control in response to carbon (C), NH4+PO4 and PO4 (one-way ANOVA, F=19.22$_{6,21}$, P<0.001; Tukey’s HSD, P=0.006, P<0.001, and P<0.001, respectively). R did not increase in response to NH4 alone, or in response to PO4 in combination with NO3, suggesting primary P limitation with some N and P co-limitation along with C limitation 

```{r}
fit <- lm(sqrt(CR.corr.no0) ~ treatment, data=limit[year=="2013" & substrate=="Sponge" & site=="Site 1"])
summary(aov(fit))
TukeyHSD(aov(fit))$treatment[c("Ctl-C", "HP-Ctl", "P-Ctl"),]
TukeyHSD(aov(fit))$treatment[c("N-Ctl", "NP-Ctl"),]
```

Reproduces

#### Site 2 and 3
> There were no significant effects of nutrient treatment on R at either the Hippo site or the Hippo + WB site, suggesting no nutrient limitation at those sites.

```{r}
fit <- lm(sqrt(CR.corr.no0) ~ treatment, data=limit[year=="2011" & substrate=="Sponge" & site=="Site 2"])
summary(aov(fit))
TukeyHSD(aov(fit))$treatment[c("H-Ctl", "P-Ctl", "N-Ctl", "HP-Ctl", "NP-Ctl"),]

fit <- lm(sqrt(CR.corr.no0) ~ treatment, data=limit[year=="2013" & substrate=="Sponge" & site=="Site 2"])
summary(aov(fit))
TukeyHSD(aov(fit))$treatment[c("H-Ctl", "P-Ctl", "N-Ctl", "HP-Ctl", "NP-Ctl"),]

```

Reproduces

```{r }
fit <- lm(sqrt(CR.corr.no0) ~ treatment, data=limit[year=="2011" & substrate=="Sponge" & site=="Site 3"])
summary(aov(fit))
TukeyHSD(aov(fit))$treatment[c("H-Ctl", "P-Ctl", "N-Ctl", "HP-Ctl", "NP-Ctl"),]

fit <- lm(sqrt(CR.corr.no0) ~ treatment, data=limit[year=="2013" & substrate=="Sponge" & site=="Site 3"])
summary(aov(fit))
TukeyHSD(aov(fit))$treatment[c("H-Ctl", "P-Ctl", "N-Ctl", "HP-Ctl", "NP-Ctl"),]

```

Reproduces

### Chlorophyll a results

>There was no significant interaction between site and treatment for chl a in either 2011 (two-way ANOVA, interaction, F=1.92$_{10,71}$, P=0.057) or 2013 (F=1.02$_{12,77}$, P=0.444).

```{r}
Chlor_hetero_Y11.fit <- lm(sqrt(Chl.corr.no0) ~ treatment*site, data=limit[year=="2011" & substrate=="Glass"])
summary(aov(Chlor_hetero_Y11.fit))

Chlor_hetero_Y13.fit <- lm(sqrt(Chl.corr.no0) ~ treatment*site, data=limit[year=="2013" & substrate=="Glass"])
summary(aov(Chlor_hetero_Y13.fit))
```

Reproduces

> At the No Wildlife site in 2011, chl a on the inorganic substrate increased significantly relative to the control in response to NH4 (one-way ANOVA, F=3.47$_{5,24}, P=0.017; Tukey’s HSD, P=0.031), suggesting some N limitation (Appendix S2: Fig. S3). There were no significant effects of nutrients on chl a at either the Hippo site or the Hippo + WB site.

```{r}
fit <- lm(sqrt(Chl.corr.no0) ~ treatment, data=limit[year=="2011" & substrate=="Glass" & site=="Site 1"])
summary(aov(fit))
TukeyHSD(aov(fit))$treatment[c("H-Ctl", "P-Ctl", "N-Ctl", "HP-Ctl", "NP-Ctl"),]

```
Reproduces

## Ecosystem Function
### Experimental Streams
#### Statistical methods

> We transformed the data to meet assumptions of normality using log transformations for GPP and R and square root transformations for AFDM and chl a. We analyzed each of the four response variables (GPP, R, AFDM and chl a) throughout the duration of the experiment using a linear mixed-effect model run with the lme function in the nlme package in R (Pinheiro et al. 2016, R Core Team 2018). We fitted lme models with the restricted maximum likelihood method and a continuous autoregressive temporal correlation structure with day as the repeated factor. Treatment (control, hippo, wb, hippo+wb) and time (each of 4 weeks of measurement) were treated as fixed effects, and individual streams were treated as random effects. A significant interaction between treatment and time indicated different rates of biofilm growth by treatment. We then used lstrends in the lsmeans package to perform a Tukey pairwise comparison test between treatments for biofilm parameters over the duration of the experiment (Lenth 2016, R Core Team 2018). We also used one-way ANOVAs to analyze differences in standing stock biofilm parameters by treatment in the final week of the experiment.

#### Results

> The rate of biofilm growth in experimental streams was influenced by the presence of wildlife inputs. There was a significant interaction between treatment and time for GPP (lme, interaction effect, F=3.03$_{3,32}$, P=0.044) 

```{r}
fit <- lme(log(GPP.no0) ~ treatment*days, random = ~ 1|id, correlation=corCAR1(form = ~ days | id), data=stream, method="REML")
anova(fit, type="marginal")

fit <- lme((GPP.no0) ~ treatment*days, random = ~ 1|id, correlation=corCAR1(form = ~ days | id), data=stream, method="REML")
anova(fit, type="marginal")

```

log(GPP) (as specified in methods) does not reproduce but untransformed GPP (or GPP.no0) does reproduce. This makes a difference if using the p-value to guide further analyses.

```{r}
fit <- lme((GPP.no0) ~ treatment*days, random = ~ 1|id, correlation=corCAR1(form = ~ days | id), data=stream, method="REML")
anova(fit, type="marginal")
```

> There was a significant interaction between treatment and time for ... chl a (lme, interaction effect, F=3.44$_{3,32}$, P=0.028) 

```{r}

fit <- lme(sqrt(Chl) ~ treatment*days, random = ~ 1|id, correlation=corCAR1(form = ~ days | id), data=stream, method="REML")
anova(fit, type="marginal")

fit <- lme((Chl) ~ treatment*days, random = ~ 1|id, correlation=corCAR1(form = ~ days | id), data=stream, method="REML")
anova(fit, type="marginal")

fit <- lme(log(Chl) ~ treatment*days, random = ~ 1|id, correlation=corCAR1(form = ~ days | id), data=stream, method="REML")
anova(fit, type="marginal")

```

Sqrt(chl) (as specified in methods) does not reproduce. Untransformed chl does not reproduce. Log(chl) does reproduce.

> The rate of increase was greatest in the Hippo + WB treatment as compared to the Control for both GPP (lstrends pairwise comparison, t ratio: -2.7632, P=0.044) and chl a (lstrends pairwise comparison, t ratio:-3.0832, P=0.021).

```{r}
# use the models that reproduce, not the models specified in methods
m1 <- lme((GPP.no0) ~ treatment*days, random = ~ 1|id, correlation=corCAR1(form = ~ days | id), data=stream, method="REML")
emtrends(m1, pairwise ~ treatment, var = "days")

m2 <- lme(log(Chl) ~ treatment*days, random = ~ 1|id, correlation=corCAR1(form = ~ days | id), data=stream, method="REML")
emtrends(m2, pairwise ~ treatment, var = "days")

```

Reproduces.

### River Biofilms
#### Statistical methods
> Data were normally distributed with equal variance among sites, so we used a one-way ANOVA followed by Tukey’s HSD test for pairwise comparisons to test for differences among sites (R Core Team 2018). We then used independent t-tests to analyze differences in chl a and AFDM (for which we had sufficient data) by site with and without the presence of wildebeest carcasses at the Hippo + WB site. There were no significant differences between seasons for any of the three sites, so we then used a one-way ANOVA followed by Tukey’s HSD test to test for differences between the No Wildlife site, the Hippo site, the Hippo + WB site with carcasses and the Hippo + WB site without carcasses.

#### Results
>GPP (n=4 per site) increased from the No Wildlife site to the Hippo site but was lowest at the Hippo + WB site (Fig. 6)

```{r}
site_levels <- c("No Wildlife", "Hippo", "Hippo + WB")
river[, site_fac:=factor(ifelse(site=="Site 1", site_levels[1],
                           ifelse(site=="Site 2", site_levels[2], site_levels[3])),
                           site_levels)
                           ]
fit <- lm(GPP ~ site_fac, data=river)
(fit.emm <- emmeans(fit, specs="site_fac"))
contrast(fit.emm, method="revpairwise")
```

reproduces qualitatively (cannot check quantitative)

> Chl a (n=6-8 per site) followed the same pattern, increasing from the No Wildlife site to the Hippo site (not significant), then decreasing from the Hippo site to the Hippo + WB site (one-way ANOVA, F=9.44$_{2,19}$, P=0.001; Tukey’s HSD: P=0.001) (Fig. 6).

```{r}
fit <- lm(Chl ~ site_fac, data=river)
anova(fit)
(fit.emm <- emmeans(fit, specs="site_fac"))
contrast(fit.emm, method="revpairwise")

```

Reproduces.

> There was also a significant difference in chl a when comparing by site and season (one-way ANOVA, F=5.99$_{3,18}$, P=0.005). The Hippo + WB site had significantly lower chl a than the Hippo site both when carcasses were present (Tukey’s HSD: P=0.015) and when they were not (Tukey’s HSD: P=0.009); however, there was no difference between the Hippo + WB site with and without carcasses (Tukey’s HSD: P=0.996).

```{r}
# river[, .(Chl=mean(Chl), N=.N), by=.(site.wb)]
# fit <- lm(Chl ~ site.wb, data=river)
# anova(fit)
# (fit.emm <- emmeans(fit, specs="site.wb"))
# contrast(fit.emm, method="revpairwise")

river[, site.wb2:=ifelse(site!="Site 3", site, site.wb)]
fit <- lm(Chl ~ site.wb2, data=river)
anova(fit)
(fit.emm <- emmeans(fit, specs="site.wb2"))
contrast(fit.emm, method="revpairwise")

```

Reproduces. However, the column "site.wb" is not coded correctly for this and I had to recode the column to reproduce.

> Both R (n=4 per site) and AFDM (n=6-7 per site) also increased from the No Wildlife site to the Hippo site (not significant). However, R and AFDM were very similar between the Hippo site and the Hippo + WB site.

```{r}
fit <- lm(CR ~ site_fac, data=river)
anova(fit)
(fit.emm <- emmeans(fit, specs="site_fac"))
contrast(fit.emm, method="revpairwise")

fit <- lm(AFDM ~ site_fac, data=river)
anova(fit)
(fit.emm <- emmeans(fit, specs="site_fac"))
contrast(fit.emm, method="revpairwise")

```

Qualitatively reproduces.

###

# Statistical red-flags

1. The reported statistics do not account for the clustered (non-independent) responses (contra the methods, which states that a repeated measures ANOVA was used).

2. Independent t-tests or t-tests using the pooled SD are different models than the original fit model with different coefficients and estimates of the original error. The recommended best practice is posthoc tests using the fit model.

3. Proportions have max variance at 0.5 and log transforms are not best practice. Instead logit or arcsine transforms are a better practice for proportions. Also, the value of adj_hippo_prop_rem for row 4 is 1.1541341, which is not allowed for proportions (so one cannot take the arcsin or logit of this)

4. The indepdendent ANOVAs are different models than the initial ANOVA with different model coefficients and estimates of the residual error (sigma). The recommended best practice is a posthoc test using the original full model.

```{r}
# original model with log transform but simulataneous pairwise comparison
m1 <- lm(ln_hippo_prop_rem ~ treatment_recode*days, data=decomp)
emtrends(m1, pairwise ~ treatment_recode, var = "days", adjust="none")
# "none" is used because the authors compare to the bonferroni 0.05/3

# what about arcsin transform? - very different
decomp[, adj_hippo_prop_rem_2:=ifelse(adj_hippo_prop_rem > 1, 1, adj_hippo_prop_rem)]
m2 <- lm(asin(adj_hippo_prop_rem_2) ~ treatment_recode*days, data=decomp)
emtrends(m2, pairwise ~ treatment_recode, var = "days", adjust="none")
# using plot() m2 is better than m1

# what about original log transform but with corrected adj_hippo_prop_rem? - not much different from original
m1b <- lm(log(adj_hippo_prop_rem_2) ~ treatment_recode*days, data=decomp)
emtrends(m1b, pairwise ~ treatment_recode, var = "days", adjust="none")

```

