---
title: 'Data from: Density-dependent positive feedbacks buffer aquatic plants from
  interactive effects of eutrophication and predator loss'
author: Jeff Walker
date: '2019-01-16'
slug: data-from-density-dependent-positive-feedbacks-buffer-aquatic-plants-from-interactive-effects-of-eutrophication-and-predator-loss
categories: []
tags:
  - excel
  - linear mixed model
  - transformed response
  - factorial
subtitle: ''
output:
  blogdown::html_page:
    toc: true
---

# Sources

Source: Donadi S, Austin ÅN, Svartgren E, Eriksson BK, Hansen JP, Eklöf JS (2018) Density-dependent positive feedbacks buffer aquatic plants from interactive effects of eutrophication and predator loss. Ecology 99(11): 2515-2524. https://doi.org/10.1002/ecy.2501

Data source: Donadi S, Austin ÅN, Svartgren E, Eriksson BK, Hansen JP, Eklöf JS (2018) Data from: Density-dependent positive feedbacks buffer aquatic plants from interactive effects of eutrophication and predator loss. Dryad Digital Repository. https://doi.org/10.5061/dryad.m1fb658

# Summary

# Caveats
Failure to reproduce or partial reproducibility might be an error in my coding, or my misunderstanding of the author's methods, or my lack of knowledge of statistics generally or the R functions that I use more specifically.

# Set up
```{r setup, message=FALSE, warning=FALSE}
library(readxl)
library(ggplot2)
library(ggpubr)
library(nlme)
library(lmerTest)
library(car)
library(emmeans)
library(data.table)
library(lmtest) #?
library(predictmeans)

data_folder <- "../data"
folder <- "Data from Density-dependent positive feedbacks buffer aquatic plants from interactive effects of eutrophication and predator loss"

# images need to be moved to /static/images/data from.../image.png
knitr_path <- FALSE

```

# Import
```{r}
fn <-"plant_grazer_data_SDonadi.xlsx"

# open sheet "shoot level"
sheet_i <- "Shoot level"
file_path <- paste(data_folder, folder, fn, sep='/')
shoot <- data.table(read_excel(file_path, sheet = sheet_i))
shoot <- shoot[CAGE_TREAT!="uncaged"] # Delete CAGE_TREAT="uncaged" as these are not in main analysis
shoot[, SITE:=factor(SITE)]
shoot[, CAGE_ID:=factor(CAGE_ID)]
shoot[, CAGE_TREAT:=factor(CAGE_TREAT)]
shoot[, DENSITY_TREAT:=factor(DENSITY_TREAT)]
shoot[, NUTRIENT_TREAT:=factor(NUTRIENT_TREAT)]

# open sheet "cage level"
sheet_i <- "Cage level"
file_path <- paste(data_folder, folder, fn, sep='/')
cage <- data.table(read_excel(file_path, sheet = sheet_i))
cage <- cage[CAGE_TREAT!="uncaged"] # Delete CAGE_TREAT="uncaged" as these are not in main analysis
cage[, SITE:=factor(SITE)]
cage[, CAGE_ID:=factor(CAGE_ID)]
cage[, CAGE_TREAT:=factor(CAGE_TREAT)]
cage[, DENSITY_TREAT:=factor(DENSITY_TREAT)]
cage[, NUTRIENT_TREAT:=factor(NUTRIENT_TREAT)]

```
## data cleaning

```{r}
shoot <- shoot[CAGE_TREAT!="uncaged"]
```

```{r, message=FALSE, warning=FALSE}
shoot[, Strictly_epiphytic_algae_DW_g_Myrio:=as.numeric(Strictly_epiphytic_algae_DW_g_Myrio)]
shoot[, scaled_Strictly_epiphytic_algae_DW_g_Myrio:=as.numeric(scaled_Strictly_epiphytic_algae_DW_g_Myrio)]
shoot[, wind_fetch:=as.numeric(wind_fetch)]


cage[, Avgscaled_macrograzers_and_chirono_AFDW_mg_Myrio:=as.numeric(Avgscaled_macrograzers_and_chirono_AFDW_mg_Myrio)]
cage[, Avgscaled_macrograzers_and_chirono_n_Myrio:=as.numeric(Avgscaled_macrograzers_and_chirono_n_Myrio)]
cage[, wind_fetch:=as.numeric(wind_fetch)]

```
```{r are all combinations present, echo=FALSE, eval=FALSE}
shoot[, .(mean=mean(ShootDW_g), N=.N), by=.(CAGE_TREAT, DENSITY_TREAT, NUTRIENT_TREAT)]
```

# Reproducibility

## Appendix S5 Table S1
Start with this table since it is the basis for the in-text results

```{r tableS1-repro, message=FALSE, warning=FALSE}
# The authors use nlme and not lme4
# fit by REML
fit1 <- lme(ShootDW_g ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, data=shoot)
fit1.anova <- anova(fit1, type = "marginal")

fit2 <- lme(sqrt(Sum_branch_length)  ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, data=shoot)
fit2.anova <- anova(fit2, type = "marginal")

fit3 <- lme(sqrt(N_tot_branches)  ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, data=shoot)
fit3.anova <- anova(fit3, type = "marginal")

fit4 <- lme(sqrt(Max_length_cm)  ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, data=shoot)
fit4.anova <- anova(fit4, type = "marginal")

fit5 <- lme(sqrt(scaled_Strictly_epiphytic_algae_DW_g_Myrio)  ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, data=shoot, na.action=na.exclude)
fit5.anova <- anova(fit5, type = "marginal")

# fit by ML
fit1b <- lme(ShootDW_g ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, method="ML", data=shoot)
fit1b.anova <- anova(fit1b, type = "marginal") # machts nichts

fit2b <- lme(sqrt(Sum_branch_length)  ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, method="ML", data=shoot)
fit2b.anova <- anova(fit2b, type = "marginal")

fit3b <- lme(sqrt(N_tot_branches)  ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, method="ML", data=shoot)
fit3b.anova <- anova(fit3b, type = "marginal")

fit4b <- lme(sqrt(Max_length_cm)  ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, method="ML", data=shoot)
fit4b.anova <- anova(fit4b, type = "marginal")

fit5b <- lme(sqrt(scaled_Strictly_epiphytic_algae_DW_g_Myrio) ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, method="ML", data=shoot, na.action=na.exclude)
fit5b.anova <- anova(fit5b, type = "marginal")

ycols <- c("F-value", "p-value")
df <- merge(fit1.anova, fit2.anova[, ycols], by=0, sort=FALSE)
df <- df[,-1]
row.names(df) <- row.names(fit1.anova)
df <- merge(df, fit3.anova[, ycols], by=0, sort=FALSE)
df <- df[,-1]
row.names(df) <- row.names(fit1.anova)
df <- merge(df, fit4.anova[, ycols], by=0, sort=FALSE)
df <- df[,-1]
row.names(df) <- row.names(fit1.anova)
df <- merge(df, fit5.anova[, ycols], by=0, sort=FALSE)
colnames(df) <- c("Term", "numDF", "denDF", "Plant biomass F", "Plant biomass p", "Branch length F", "Branch length p", "N branches F", "N branches p", "Stem length F", "Stem length p", "Algae biomass F", "Algae biomass p")
table_s1_reml <- copy(df)
knitr::kable(table_s1_reml, digits=c(NA, 0, 0, 2, 3, 2, 3, 2, 3, 2, 3, 2, 3))

ycols <- c("F-value", "p-value")
df <- merge(fit1b.anova, fit2b.anova[, ycols], by=0, sort=FALSE)
df <- df[,-1]
row.names(df) <- row.names(fit1b.anova)
df <- merge(df, fit3b.anova[, ycols], by=0, sort=FALSE)
df <- df[,-1]
row.names(df) <- row.names(fit1b.anova)
df <- merge(df, fit4b.anova[, ycols], by=0, sort=FALSE)
df <- df[,-1]
row.names(df) <- row.names(fit1b.anova)
df <- merge(df, fit5b.anova[, ycols], by=0, sort=FALSE)
colnames(df) <- c("Term", "numDF", "denDF", "Plant biomass F", "Plant biomass p", "Branch length F", "Branch length p", "N branches F", "N branches p", "Stem length F", "Stem length p", "Algae biomass F", "Algae biomass p")
table_s1_ml <- copy(df)
knitr::kable(table_s1_ml, digits=c(NA, 0, 0, 2, 3, 2, 3, 2, 3, 2, 3, 2, 3))

```

```{r tableS1, echo=FALSE}
fn <- "tableS1.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

### reproducing epiphytic algae biomass columns

An additive model for epiphyte biomass reproduces the first 6 rows of Table S1 (the last 4 are from the full factorial model fitted by ML). Found this not by trying to reproduce table but because I wanted to compare p-values of marginal effects with and without interactions in the model.

```{r tableS1-explore}
# 
fit5c <- lme(sqrt(scaled_Strictly_epiphytic_algae_DW_g_Myrio) ~ DENSITY_TREAT + CAGE_TREAT + NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE/CAGE_ID, method="ML", data=shoot, na.action=na.exclude)
knitr::kable(anova(fit5c, type = "marginal"), digits=c(0,0,2,3))

```

1. branch length and stem length reproduce using REML

2. plant biomass and number of branches reproduce using ML

3. The main effects for epiphyte algae biomass reproduces using an additive model (no interactions) and ML while the interaction effects reproduce using the full factorial model and ML. I didn't see in the text that the authors used an additive model and the table suggests the full factorial model.

No information in methods why the authors use REML for some responses and ML for others.

## in-text

>After 12 weeks, epiphytic algae biomass was 2.4 times higher in the large predatory fish exclosures (closed cages) than the fish access (open) cages but was unaffected by nutrient enrichment and plant density (main effect of cage treatment, F1,19 = 6.12, P = 0.020, Appendix S5, Fig. 2A)

-- The ANOVA F and p reproduce using an additive model (the methods and Table S1 imply interactions included) and ML estimation. See my Table S1 for full table. The reported ddf differs from my table because the reported ddf report the ddf "as if" the full-factorial model were used.

```{r}
# additive  model
knitr::kable(anova(fit5c, type = "marginal")["CAGE_TREAT",], digits=c(0,0,2,3))
```

```{r}
# algae mass used the additive model
fit.emm <- data.table(summary(emmeans(fit5c, specs="CAGE_TREAT"))) # marginal means
a <- (fit.emm[CAGE_TREAT=="closed", emmean])^2
b <- (fit.emm[CAGE_TREAT=="open", emmean])^2
a/b
```

This doesn't reproduce. Here I try the raw means that do not account for clustered data and unbalanced design.

```{r}
a <- mean(shoot[CAGE_TREAT=="closed", scaled_Strictly_epiphytic_algae_DW_g_Myrio], na.rm=TRUE)
b <- mean(shoot[CAGE_TREAT=="open", scaled_Strictly_epiphytic_algae_DW_g_Myrio], na.rm=TRUE)
a/b

```

This reproduces. See section "Statistical Red-flag" #4

>Plant density interacted with nutrient enrichment and exclusion of large fish (three-way interaction, Appendix S5) to affect plant performance: in closed cages (where algal bio- mass was generally higher, see above) and under nutrient- enriched conditions, high plant density increased individual plant biomass 4.7 times compared to medium plant density (Fisher’s LSD test, F2,19 = 6.19, P = 0.009, Fig. 3A)

I assumed that this p value was that for the post hoc comparison but it (and the F ratio) is from the row for the 3-way interaction in the ANOVA table, which I reproduce (see Table S1 below). However, this is not the row in the ANOVA table used to justify or "protect" posthoc tests within a factor (like Density). For the actual p-value of this contrast, see "statistical red flags #3" below.

I assume that "increase 4.7X" is the fraction increase. But this doesn't reproduce.

```{r}


# is the fold-change correct?
a <- data.table(summary(fit.emm))[DENSITY_TREAT=="3_high" & CAGE_TREAT=="closed" & NUTRIENT_TREAT=="enriched", emmean]
b <- data.table(summary(fit.emm))[DENSITY_TREAT=="2_intermediate" & CAGE_TREAT=="closed" & NUTRIENT_TREAT=="enriched", emmean]
(a-b)/b
```

Here I used the raw means (ignoring the clustered structure) instead of the modeled means, and it does reproduce. 
```{r}
a <- mean(shoot[DENSITY_TREAT=="3_high" & CAGE_TREAT=="closed" & NUTRIENT_TREAT=="enriched", ShootDW_g], na.rm=TRUE)
b <- mean(shoot[DENSITY_TREAT=="2_intermediate" & CAGE_TREAT=="closed" & NUTRIENT_TREAT=="enriched", ShootDW_g], na.rm=TRUE)
(a-b)/b
a/b
```

This reproduces if I use the raw means that do not account for the clustered data and unbalanced design and if the value is not the "increase" but simply the multiple (it didn't increase 4.7X but is simply 4.7X bigger). See section "Statistical Red-flag" #4

> in closed cages ... under nutrient-enriched conditions [and] high plant density ... number of branches per plant 2.3 and 2.8 times (Fisher’s LSD test, F2,19 = 3.74, P = 0.043, Fig. 3B), compared to medium and low plant density, respectively. 

I assume the multiple is using raw data as above. Again, the ANOVA stats are for the 3-way interaction row in TABLE S1 and this p-value is not the p-value of the contrasts. These p-vales are in "Statistical Red-flag" #5

## Appendix S5 Table S2

```{r}
# fit by REML
fit6 <- lme(Avgscaled_macrograzers_and_chirono_AFDW_mg_Myrio ~ DENSITY_TREAT*CAGE_TREAT*NUTRIENT_TREAT + wind_fetch + Actual_Retention_time, random = ~1|SITE, method="REML", data=cage, na.action=na.exclude)
anova(fit6, type = "marginal")
fit6 <- lme(Avgscaled_macrograzers_and_chirono_AFDW_mg_Myrio ~ DENSITY_TREAT + CAGE_TREAT + NUTRIENT_TREAT + wind_fetch + Actual_Retention_time + DENSITY_TREAT:CAGE_TREAT + DENSITY_TREAT:NUTRIENT_TREAT + CAGE_TREAT*NUTRIENT_TREAT + DENSITY_TREAT:CAGE_TREAT:NUTRIENT_TREAT, random = ~1|SITE, method="ML", data=cage, na.action=na.exclude)
# - Density Treat
fit6a <- lme(Avgscaled_macrograzers_and_chirono_AFDW_mg_Myrio ~ CAGE_TREAT + NUTRIENT_TREAT + wind_fetch + Actual_Retention_time + DENSITY_TREAT:CAGE_TREAT + DENSITY_TREAT:NUTRIENT_TREAT + CAGE_TREAT*NUTRIENT_TREAT + DENSITY_TREAT:CAGE_TREAT:NUTRIENT_TREAT, random = ~1|SITE, method="ML", data=cage, na.action=na.exclude)
anova(fit6a, type = "marginal")
anova(fit6a, fit6, test=TRUE)
lrtest(fit6, fit6a)

```

## Figure 2

```{r}
# raw SE
shoot[, present_scaled_Strictly_epiphytic_algae_DW_g_Myrio:=ifelse(is.na(scaled_Strictly_epiphytic_algae_DW_g_Myrio),0,1)]
fit5.raw_means <- shoot[, .(epiph_algae_biom=mean(scaled_Strictly_epiphytic_algae_DW_g_Myrio, na.rm=TRUE),
                            sd=sd(scaled_Strictly_epiphytic_algae_DW_g_Myrio, na.rm=TRUE),
                            N=sum(present_scaled_Strictly_epiphytic_algae_DW_g_Myrio)),
                        by=.(NUTRIENT_TREAT, CAGE_TREAT, DENSITY_TREAT)]
fit5.raw_means[, SE:=sd/sqrt(N)]
# orderBy(~NUTRIENT_TREAT + CAGE_TREAT + DENSITY_TREAT, fit5.raw_means)

pd <- position_dodge(0.9)
gg1 <- ggplot(data=fit5.raw_means, aes(x=CAGE_TREAT, y=epiph_algae_biom, fill=DENSITY_TREAT)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=epiph_algae_biom-SE, ymax=epiph_algae_biom+SE), width=0.2, position=pd) +
  facet_wrap(.~NUTRIENT_TREAT) +
  NULL
gg1
```

# Statistical Red-flags

1. Two different methods (REML and ML) used for fitting models in the main ANOVA table (Appendix 5, Table S1)

2. The ANOVA for algae biomass in Appendix 5, Table S1 is composite -- the main effects is the ANOVA from an additive model while the interactions are from the full-factorial model.

3. The F-ratio/p value for the 3-way interaction in Table S1 is not the p value used to "protect" post hoc comparisons among levels of Density within some combination of Nutrients and Cage. With this complex of a design I don't think there is any row of the ANOVA table that "protects" post-hoc comparisons. If one wants unadjusted p-values for a specific post-hoc comparison (I'm fine with this), I would use the contrast function of the emmeans package.

```{r}
# plant biomass ANOVA reproduced with ML so use fit1b
fit.emm <- emmeans(fit1b, specs=c("DENSITY_TREAT", "CAGE_TREAT", "NUTRIENT_TREAT"))
fit.contrast <- data.table(summary(contrast(fit.emm, method="revpairwise", adjust="none")))[contrast=="3_high,closed,enriched - 2_intermediate,closed,enriched",]
fit.contrast$p.value

fit.pm <- predictmeans(fit1b, modelterm="DENSITY_TREAT:CAGE_TREAT:NUTRIENT_TREAT", pairwise = TRUE, plot=FALSE)
fit.pm$"Pairwise p-value"["3_high:closed:enriched","2_intermediate:closed:enriched"]
```

I'm not sure why these differ, I would think they would be the same but its not clear to me how the "predictmeans" LSD pairwise comparison is computing t-values

4. The fold-changes given in the text are based on the means computed from the raw data and not from the model. This can sometimes be misleading. See https://cran.r-project.org/web/packages/emmeans/vignettes/basics.html

5. p-values for number of branches

Again, not sure why these differ, I would think they would be the same. Regardless these are the p-values that should be reported with a contrast.

```{r}
# number of branches reproduced with ML so use fit3b
fit.emm <- emmeans(fit3b, specs=c("DENSITY_TREAT", "CAGE_TREAT", "NUTRIENT_TREAT"))
fit.contrast <- data.table(summary(contrast(fit.emm, method="revpairwise", adjust="none")))[contrast=="3_high,closed,enriched - 2_intermediate,closed,enriched",]
fit.contrast$p.value

fit.pm <- predictmeans(fit3b, modelterm="DENSITY_TREAT:CAGE_TREAT:NUTRIENT_TREAT", pairwise = TRUE, plot=FALSE)
fit.pm$"Pairwise p-value"["3_high:closed:enriched","2_intermediate:closed:enriched"]

```

## Figure red-flags

The SE from the figures are based on raw data and do not represent the model and cannot be used to infer anything useful because 1) they do not take into account the random effects and they are taken from untransformed (not sqrt transformed) data. Note how 2*SE for many of the error bars would extend (often well) below zero.

Figures that better represent the model (the last doesn't, but I use it to get "raw" or "unmodeled" SE that ignore random effects)

```{r}
# modeled SE using the full-factorial model and ML fit (not the additive model reported)
fit5b.emm <- data.table(summary(emmeans(fit5b, 
              specs=c("DENSITY_TREAT", "CAGE_TREAT","NUTRIENT_TREAT"))))
fit5b.emm[, emmean.sq:=emmean^2]
fit5b.emm[, SE.sq:=SE^2]

# modeled SE using the additve model and ML fit
fit5c.emm <- data.table(summary(emmeans(fit5c, 
              specs=c("DENSITY_TREAT", "CAGE_TREAT","NUTRIENT_TREAT"))))
fit5c.emm[, emmean.sq:=emmean^2]
fit5c.emm[, SE.sq:=SE^2]

# raw SE using transformed data then backtransformed
fit5.raw_means <- shoot[, .(epiph_algae_biom=mean(sqrt(scaled_Strictly_epiphytic_algae_DW_g_Myrio), na.rm=TRUE),
                            sd=sd(sqrt(scaled_Strictly_epiphytic_algae_DW_g_Myrio), na.rm=TRUE),
                            N=sum(present_scaled_Strictly_epiphytic_algae_DW_g_Myrio)),
                        by=.(NUTRIENT_TREAT, CAGE_TREAT, DENSITY_TREAT)]
fit5.raw_means[, SE:=sd/sqrt(N)]
fit5.raw_means[, raw_mean:=epiph_algae_biom^2]
fit5.raw_means[, raw_SE:=SE^2]

# full factorial model
pd <- position_dodge(0.9)
gg1b <- ggplot(data=fit5b.emm, aes(x=CAGE_TREAT, y=emmean.sq, fill=DENSITY_TREAT)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=emmean.sq-SE.sq, ymax=emmean.sq+SE.sq), width=0.2, position=pd) +
  facet_wrap(.~NUTRIENT_TREAT) +
  NULL
gg1b

# additive model
pd <- position_dodge(0.9)
gg1c <- ggplot(data=fit5c.emm, aes(x=CAGE_TREAT, y=emmean.sq, fill=DENSITY_TREAT)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=emmean.sq-SE.sq, ymax=emmean.sq+SE.sq), width=0.2, position=pd) +
  facet_wrap(.~NUTRIENT_TREAT) +
  NULL
gg1c

# raw backtransformed - SE reflect "pseudoreplication"
pd <- position_dodge(0.9)
gg1d <- ggplot(data=fit5.raw_means, aes(x=CAGE_TREAT, y=raw_mean, fill=DENSITY_TREAT)) +
  geom_col(position=pd) +
  geom_errorbar(aes(ymin=raw_mean-raw_SE, ymax=raw_mean+raw_SE), width=0.2, position=pd) +
  facet_wrap(.~NUTRIENT_TREAT) +
  NULL
gg1d

```

