---
title: 'Data from: Litter removal in a tropical rain forest reduces fine root biomass
  and production but litter addition has few effects'
author: Jeff Walker
date: '2019-01-14'
slug: data-from-litter-removal-in-a-tropical-rain-forest-reduces-fine-root-biomass-and-production-but-litter-addition-has-few-effects
categories: []
tags: []
subtitle: ''
output:
  blogdown::html_page:
    toc: true
---

# Sources

Rodtassana C, Tanner EVJ (2018) Litter removal in a tropical rain forest reduces fine root biomass and production but litter addition has few effects. Ecology 99(3): 735-742. https://doi.org/10.1002/ecy.2143

Rodtassana C, Tanner EVJ (2018) Data from: Litter removal in a tropical rain forest reduces fine root biomass and production but litter addition has few effects. Dryad Digital Repository. https://doi.org/10.5061/dryad.f1948

# Summary

# Set up

```{r setup, warning=FALSE, message=FALSE}
library(readxl)
library(ggplot2)
library(ggpubr)
library(cowplot)
library(lmerTest)
library(emmeans)
library(data.table)
library(multcompView) # grouping post hoc (not recommended!)


data_folder <- "../data"
folder <- "Data from Litter removal in a tropical rain forest reduces fine root biomass and production but litter addition has few effects"

# images need to be moved to /static/images/data from.../image.png
knitr_path <- FALSE
```

## Functions

Some code used for model selection with a mixed model following Chapter 5.7 of Zuur for model selection of mixed effect models.

```{r functions}

select_model <- function(y, x, r, dt, reml=FALSE){
  # the data are in dt
  # y is the response column name
  # x is the list of fixed effect column names
  # r is the list of random effect column names
  # it would seem weird to have both month and plot in model
  # see 5.7 of Zuur for model selection of mixed effect models
  fixed <- paste(y,"~", paste(x, collapse='*'))
  random <- list()
  random[[1]] <- paste(paste("(1|", r, ")", sep=""), collapse=" + ")
  min_aic <- 1
  
  # fit random part first, if there are > 1 random variables
  if(length(r) > 1){
    aic <- numeric(3)
    form <- formula(paste(fixed, random[[1]], sep=' + '))
    fit1 <- lmer(form, data=dt, REML=TRUE)
    aic[1] <- AIC(fit1)
    r1 <- r[1] # just the first random element
    random[[2]] <- paste(paste("(1|", r1, ")", sep=""), collapse=" + ")
    form <- formula(paste(fixed, random[[2]], sep=' + '))
    fit2 <- lmer(form, data=dt, REML=TRUE)
    aic[2] <- AIC(fit2)
    r2 <- r[2] # just the second random element
    random[[3]] <- paste(paste("(1|", r2, ")", sep=""), collapse=" + ")
    form <- formula(paste(fixed, random[[3]], sep=' + '))
    fit3 <- lmer(form, data=dt, REML=TRUE)
    aic[3] <- AIC(fit3)
    min_aic <- which(aic == min(aic))
  }
  random <- random[[min_aic]]

  # fit fixed part
  form <- list()
  # fit with ML, choose model, then refit with REML
  aic <- numeric(4)
  form[[1]] <- formula(paste(fixed, random, sep=' + '))
  # default is REML
  fit1 <- lmer(form[[1]], data=dt, REML=reml)
  aic[1] <- AIC(fit1)
  
  fixed <- paste(y,"~", paste(x, collapse=' + '))
  form[[2]] <- formula(paste(fixed, random, sep=' + '))
  fit2 <- lmer(form[[2]], data=dt, REML=reml)
  aic[2] <- AIC(fit2)

  fixed <- paste(y,"~", x[1])
  form[[3]] <- formula(paste(fixed, random, sep=' + '))
  fit3 <- lmer(form[[3]], data=dt, REML=reml)
  aic[3] <- AIC(fit3)

  fixed <- paste(y,"~", x[2])
  form[[4]] <- formula(paste(fixed, random, sep=' + '))
  fit4 <- lmer(form[[4]], data=dt, REML=reml)
  aic[4] <- AIC(fit4)

  min_aic <- which(aic == min(aic))
  return(form[[min_aic]])
}

```

# import
```{r import}
fn <- 'Root_Ecology_supplementary_materials.xlsx'
file_path <- paste(data_folder, folder, fn, sep='/')
root <- data.table(read_excel(file_path, range = cell_rows(1:16)))
```

## Note in file

Note in Excel data sheet
> FRB = Fine root biomass (g/m2) at 0-5 cm and 0-10 cm
FRL = Fine root length (m/m2 of root window) at 0-5 cm and 0-10 cm
Sur = Fine root length survivorship (m/m2 of root window) at 0-10 cm
Pro =Fine root length production (m/m2 of root window) at 0-10 cm
IC = Fine root production from ingrowth cores (g/m2/3-months)

## Data inconsistencies
Within the FRL columns are a column labeled FRB5cmJul13 (Excel column "AB") and a column labeled FRB10cmJul13 (Excel column "AN"). I am treating these ase FRL and not FRB

## Data pre-processing
data are in wide format -- reshape to long. Also need to create "season" factor with levels ("wet", "dry")

1. "There were three separate sets of ingrowth cores set up in different seasons and collected three months after installation: wet season (May to August 2013), transition period (October 2013 to January 2014), and dry season (January to April 2014)."
2. "Composite samples were made for each of six months – four in the wet season in June, September, October and December in 2013 and two in the dry season in March 2013 and February 2014."
3. "We generated several models composed of different fixed factors as litter treatments (litter removal, litter addition, control), seasons (wet and dry; transition period only for ingrowth cores) and their interactions (treatment x season);"
4. Fig S1 explicitly defines May-Dec as "wet"

```{r fine root biomass subset}
# root is in wide format - reshape to long
FRB_cols <- c("FRB5cmMar13", "FRB5cmApr13", "FRB5cmMay13", "FRB5cmJun13", "FRB5cmJul13", "FRB5cmAug13", "FRB5cmSep13", "FRB5cmOct13", "FRB5cmNov13", "FRB5cmDec13", "FRB5cmJan14", "FRB5cmFeb14", "FRB10cmMar13", "FRB10cmApr13", "FRB10cmMay13", "FRB10cmJun13", "FRB10cmJul13", "FRB10cmAug13", "FRB10cmSep13", "FRB10cmOct13", "FRB10cmNov13", "FRB10cmDec13")
# what are these "FRB5cmJul13__1", "FRB10cmJul13__1" they have values more like FRL than FRB and physically they are within the FRL set of columns (in the exact place of FRL5cmJul and FRL10cmJul)
frb <- melt(root, id.vars=c('treatment', 'plot'), measure.vars=FRB_cols, variable.name='label', value.name='FRB')
frb[, treatment:=factor(treatment)] #CT-control; LA-litter addition; LR-litter removal
frb[, label:=as.character(label)]
frb[, depth:=factor(ifelse(substr(label, 4, 4)=='5', 5, 10))] # depth is 4th character in label
frb[, month:=substr(label, nchar(label)-4, nchar(label)-2)] # month is in label
dry <- c('Jan', 'Feb', 'Mar', 'Apr') # wet is everything else
# May-Dec is "wet" per fig S1
frb[, season:=factor(ifelse(month %in% dry, 'dry', 'wet'))]

```

```{r fine root biomass production}
ic_cols <- c("ICAug13-Oct13", "ICOct13-Jan14", "ICJan14-Apr14")
frb_prod <- melt(root, id.vars=c('treatment', 'plot'), measure.vars=ic_cols, variable.name='label', value.name='frb_prod')
frb_prod[, season:=factor(ifelse(label=="ICAug13-Oct13", "wet", 
                            ifelse(label=="ICOct13-Jan14", "transitional", "dry")),
                          c("wet", "transitional", "dry"))]
```

For FRL the "FRB5cmJul13" and "FRB10cmJul13" are (probably) FRL and not FRB and imported as "FRB5cmJul13__1" and "FRB10cmJul13__1"
```{r fine root length subset}
root[, FRL5cmJul13:=FRB5cmJul13__1]
root[, FRL10cmJul13:=FRB10cmJul13__1]

FRL_cols <- c("FRL5cmJun13",	"FRL5cmJul13",	"FRL5cmAug13",	"FRL5cmSep13",	"FRL5cmOct13",	"FRL5cmNov13",	"FRL5cmDec13",	"FRL5cmJan14",	"FRL5cmFeb14",	"FRL5cmMar14",	"FRL5cmApr14",	"FRL5cmMay14",	"FRL10cmJun13",	"FRL10cmJul13",	"FRL10cmAug13",	"FRL10cmSep13",	"FRL10cmOct13",	"FRL10cmNov13",	"FRL10cmDec13",	"FRL10cmJan14",	"FRL10cmFeb14",	"FRL10cmMar14",	"FRL10cmApr14",	"FRL10cmMay14")
frl <- melt(root, id.vars=c('treatment', 'plot'), measure.vars=FRL_cols, variable.name='label', value.name='FRL')
frl[, treatment:=factor(treatment)] #CT-control; LA-litter addition; LR-litter removal
frl[, label:=as.character(label)]
frl[, depth:=factor(ifelse(substr(label, 4, 4)=='5', 5, 10))] # depth is 4th character in label
frl[, month:=substr(label, nchar(label)-4, nchar(label)-2)] # month is in label
dry <- c('Jan', 'Feb', 'Mar', 'Apr') # wet is everything else
# May-Dec is "wet" per fig S1
frl[, season:=factor(ifelse(month %in% dry, 'dry', 'wet'))]

```

```{r fine root length survival subset}
Sur_cols <- c("SurJul13", "SurAug13",	"SurSep13",	"SurOct13",	"SurNov13",	"SurDec13",	"SurJan14",	"SurFeb14",	"SurMar14",	"SurApr14",	"SurMay14")
frl_surv <- melt(root, id.vars=c('treatment', 'plot'), measure.vars=Sur_cols, variable.name='label', value.name='FRL_surv')
frl_surv[, treatment:=factor(treatment)] #CT-control; LA-litter addition; LR-litter removal
frl_surv[, label:=as.character(label)]
frl_surv[, month:=substr(label, nchar(label)-4, nchar(label)-2)] # month is in label
dry <- c('Jan', 'Feb', 'Mar', 'Apr') # wet is everything else
# May-Dec is "wet" per fig S1
frl_surv[, season:=factor(ifelse(month %in% dry, 'dry', 'wet'))]
```

```{r fine root length production subset}
pro_cols <- c("ProJul13",	"ProAug13",	"ProSep13",	"ProOct13",	"ProNov13",	"ProDec13",	"ProJan14",	"ProFeb14",	"ProMar14",	"ProApr14",	"ProMay14")
frl_prod <- melt(root, id.vars=c('treatment', 'plot'), measure.vars=pro_cols, variable.name='label', value.name='FRL_prod')
frl_prod[, treatment:=factor(treatment)] #CT-control; LA-litter addition; LR-litter removal
frl_prod[, label:=as.character(label)]
frl_prod[, month:=substr(label, nchar(label)-4, nchar(label)-2)] # month is in label
dry <- c('Jan', 'Feb', 'Mar', 'Apr') # wet is everything else
# May-Dec is "wet" per fig S1
frl_prod[, season:=factor(ifelse(month %in% dry, 'dry', 'wet'))]
```

# Reproducibility

## Section: Fine root mass and production

"Fine root mass in the soil was higher in the wet season than in the dry season, both at 0-5 cm soil depth (F1, 178 = 10.2, P<0.01, Fig. 1) and at 0-10 cm soil depth (F1, 178 = 14.9, P<0.001, Fig. 1), but there were no inter-seasonal differences in the fine root mass in the litter standing crop. Fine root mass production, over 0-10 cm soil depth, was higher in the wet season than in the transition (wet to dry) and dry seasons (F2, 87 = 52.6, P<0.001, Fig. 2)."


```{r}
form <- select_model(y="FRB", x=c('treatment', 'season'), r=c('plot', 'month'), frb[depth=="5"], reml=FALSE)
# FRB ~ treatment + season + (1 | plot) + (1 | month)
frb5.fit <- lmer(form, REML=TRUE, data=frb[depth=="5"])
frb5.anova <- anova(frb5.fit, type=3)

form <- select_model(y="FRB", x=c('treatment', 'season'), r=c('plot', 'month'), frb[depth=="10"], reml=FALSE)
# FRB ~ season + (1 | plot) + (1 | month)
frb10.anova <- anova(lmer(form, REML=TRUE, data=frb[depth=="10"]), type=3)

my.anova.table <- data.table(response="FRB", 
                             depth=5, 
                             NDF=frb5.anova["season", "NumDF"],
                             DDF=frb5.anova["season", "DenDF"],
                             F=frb5.anova["season", "F value"],
                             p=frb5.anova["season", "Pr(>F)"])
my.anova.table <- rbind(my.anova.table,
                        data.table(response="FRB", 
                             depth=10, 
                             NDF=frb10.anova["season", "NumDF"],
                             DDF=frb10.anova["season", "DenDF"],
                             F=frb10.anova["season", "F value"],
                             p=frb10.anova["season", "Pr(>F)"]))

```

### My ANOVA results
```{r}
my.anova.table
```

### reported ANOVA results
```{r echo=FALSE}
#(F1, 178 = 10.2, P<0.01, Fig. 1) and at 0-10 cm soil depth (F1, 178 = 14.9, P<0.001, Fig. 1)
their.anova.table <- data.table(response="FRB", 
                             depth=5, 
                             NDF=1,
                             DDF=178,
                             F=10.2,
                             p="P<0.01")
their.anova.table <- rbind(their.anova.table,
                        data.table(response="FRB", 
                             depth=10, 
                             NDF=1,
                             DDF=178,
                             F=14.9,
                             p="P<0.001"))
their.anova.table
```

My F for 5 cm depth is the same (to rounding) as that reported but the DDF for the reported are 178, which would be the case with a simple one-way ANOVA with no treatment effect and no random effects. The authors do not report the fit model and so it is hard to compare.

MY F for 10 cm depth is not close to the reported F. Again, the authors do not report the fit model. The DDF for the simple one way model is only 148 but the authors report 178 for ddf.

### Did the authors use a one way ANOVA with no random effects?
No. DDF are the same for 5cm but not 10 cm and the F-ratios are not close.
```{r}
anova(lm(FRB ~ season, data=frb[depth=="5"]))
anova(lm(FRB ~ season, data=frb[depth=="10"]))
```

### Figure 1

Hard to know what SEM is reported in Figure 1. These could be
1. SEM of the raw data, ignoring clustering effect of plot. This isn't too useful unless its corrected for correlation.
2. SEM of the plot means averaged over months. This is not right because it could just as well be the SEM of the month means averaged over plots.
3. modeled SEM from the linear mixed model. I prefer this.

Given the variation in SEM bars in the reported fig. 1, I am assuming that the reported SEM is not #3. So, I compute both #1 and #2 here. It looks like the reported SEM are something like #2.

```{r figure1-repro}
#1
frb5_se1 <- frb[depth==5,.(N=.N,
              FRB=mean(FRB),
              SE=sd(FRB)/sqrt(.N)),
              by=.(treatment, season)]
frb5.p1 <- frb[depth==5, .(FRB=mean(FRB)),
              by=.(treatment, season, plot)]
frb5_se2 <- frb5.p1[,.(N=.N,
              FRB=mean(FRB),
              SE=sd(FRB)/sqrt(.N)),
              by=.(treatment, season)]
frb5_se3 <- summary(emmeans(frb5.fit, specs=c("treatment", "season")))
frb5 <- frb5_se2

frb10_se1 <- frb[depth==10, .(N=.N,
              FRB=mean(FRB),
              SE=sd(FRB)/sqrt(.N)),
              by=.(treatment, season)]
frb10.p1 <- frb[depth==10, .(FRB=mean(FRB)),
              by=.(treatment, season, plot)]
frb10_se2 <- frb10.p1[,.(N=.N,
              FRB=mean(FRB),
              SE=sd(FRB)/sqrt(.N)),
              by=.(treatment, season)]
#frb10_se3 <- summary(emmeans(frb10.fit, specs=c("treatment", "season")))
frb10 <- frb10_se2

frb_sum <- rbind(data.table(depth=5, frb5), data.table(depth=10, frb10))
frb_sum[, Lower:=FRB-SE]
frb_sum[, Upper:=FRB+SE]
frb_sum[, season:=factor(season, c('wet', 'dry'))]
frb_sum[, treatment:=factor(treatment, c('LR', 'CT', 'LA'))]
pd <- position_dodge(.6)
gg <- ggplot(data=frb_sum, aes(x=season, y=FRB, fill=treatment)) +
  geom_col(position=pd, width=0.6, color='black') +
  geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2, position=pd) +
  scale_fill_manual(values=c("white", "gray", "black")) +
  facet_grid(.~depth, labeller = label_both) +
  theme_minimal(base_size=14)
gg
```

```{r fig1, echo=FALSE}
fn <- "fig1.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)
```

These look about right but I'm not sure where the "litter standing crop" biomass data are.

### post-hoc tests for fig 1
Treatment is not in the model for 10cm subset so this is only for the 5 cm data. This reproduces the reported result (cannot compare quantitatively)

```{r}
fit5.emm <- emmeans(frb5.fit, specs=c("treatment"))
contrast(fit5.emm, method="revpairwise")
CLD(fit5.emm) # not recommended as this is classification based on arbitrary boundaries
```

### Fine root mass production
```{r}
form <- select_model(y="frb_prod", x=c('treatment', 'season'), r=c('plot'), frb_prod, reml=FALSE)
# frb_prod ~ season + (1 | plot)
frb_prod.fit <- lmer(form, REML=FALSE, data=frb_prod)
frb_prod.emm <- emmeans(frb_prod.fit, specs="season")
frb_prod.anova <- anova(frb_prod.fit, type=3)
frb_prod.anova
```

reported: "F2, 87 = 52.6, P<0.001"

Again my DDF and F are not close to the reported values. There are only 3 "IC" columns. If these are the correct columns then there are 3*15=45 total points so there cannot be 87 ddf.

### effects of treatment on FRB and FRB production

1. "Fine root mass was lower in litter removal soils at 0-5 cm (significant, Fig. 1), and 0-10 cm soils (not significant, Fig. 1 and Appendix S1: Fig. S1)" -- reproduces qualitatively, see my table of contrasts and my figure 1.
2. "Litter addition did not significantly lower fine root mass in either 0-5 cm or 0-10 cm soils" - reproduces qualitatively - see figure 1
3. "The sum of the fine root mass in the litter standing crop and 0-5 cm soils was significantly lower in litter removal than controls (F2, 177 = 7.0, P<0.01)" -- cannot check because there is no apparent litter standing crop mass
4. "Fine root mass production (in 0-10 cm soil) was not significantly affected by litter removal or litter addition (Fig. 2)." -- treatment was removed from my model using the model selection algorithm. But here are the effects of treatment on FRB-prod (rows "treatmentLA" and "treatmentLR")

```{r}
coef(summary(lmer(frb_prod ~ treatment + season + (1 | plot), REML=TRUE, data=frb_prod)))
```

### figure 2
```{r fig2-repro}
# no repeated measures within treatment x season
frb_prod.se <- frb_prod[, .(N=.N,
              frb_prod=mean(frb_prod),
              SE=sd(frb_prod)/sqrt(.N)),
              by=.(treatment, season)]
frb_prod.se[, treatment:=factor(treatment, c("LR", "CT", "LA"))]
pd <- position_dodge(.6)
gg <- ggplot(data=frb_prod.se, aes(x=season, y=frb_prod, fill=treatment)) +
  geom_col(position=pd, width=0.6, color='black') +
  geom_errorbar(aes(ymin=frb_prod-SE, ymax=frb_prod+SE), width=.2, position=pd) +
  scale_fill_manual(values=c("white", "gray", "black")) +
  theme_minimal(base_size=14)
gg
```

```{r fig2, echo=FALSE}
fn <- "fig2.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

## Section: Fine root length: standing crop, production and survivorship
### table 1
```{r frl 5}
form <- select_model(y="FRL", x=c('treatment', 'season'), r=c('plot', 'month'), frl[depth=="5"], reml=FALSE)
# FRL ~ treatment * season + (1 | plot) + (1 | month)
frl5.fit <- lmer(form, REML=TRUE, data=frl[depth=="5"])
frl5.anova <- anova(frl5.fit, type=3)
```

```{r frl 10}
form <- select_model(y="FRL", x=c('treatment', 'season'), r=c('plot', 'month'), frl[depth=="10"], reml=FALSE)
# FRL ~ treatment * season + (1 | plot) + (1 | month)
frl10.fit <- lmer(form, REML=TRUE, data=frl[depth=="10"])
frl10.anova <- anova(frl10.fit, type=3)
```

```{r frl production}
form <- select_model(y="FRL_prod", x=c('treatment', 'season'), r=c('plot', 'month'), frl_prod, reml=FALSE)
# FRL_prod ~ treatment * season + (1 | plot) + (1 | month)
frl_prod.fit <- lmer(form, REML=TRUE, data=frl_prod)
frl_prod.anova <- anova(frl_prod.fit, type=3)
```

```{r frl survival}
form <- select_model(y="FRL_surv", x=c('treatment', 'season'), r=c('plot', 'month'), frl_surv, reml=FALSE)
# treatment * season + (1 | plot) + (1 | month)
frl_surv.fit <- lmer(form, REML=TRUE, data=frl_surv)
frl_surv.anova <- anova(frl_surv.fit, type=3)

```
#### my table
fine root length 5 cm depth
```{r}
knitr::kable(frl5.anova[, c("NumDF", "F value", "Pr(>F)")], digits=c(1, 2, 4))
```

fine root length 10 cm depth
```{r}
knitr::kable(frl10.anova[, c("NumDF", "F value", "Pr(>F)")], digits=c(1, 2, 4))
```

Fine root length production
```{r}
knitr::kable(frl_prod.anova[, c("NumDF", "F value", "Pr(>F)")], digits=c(1, 2, 4))
```

Fine root length survival
```{r}
knitr::kable(frl_surv.anova[, c("NumDF", "F value", "Pr(>F)")], digits=c(1, 2, 4))
```

#### reported table
```{r table1, echo=FALSE}
fn <- "table1.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)

```

The F ratios have the same basic pattern but do not reproduce

#### Alternative specifications to recover authors' F-ratios
none do.

```{r echo=FALSE}
anova(lmer(FRL ~ treatment * season + (1 | plot) + (1 | month), data=frl[depth=="5"]), type="3")
anova(lmer(FRL ~ treatment * season + (1 | plot) + (1 | month), data=frl[depth=="5"]), type="2")

anova(lmer(FRL ~ treatment * season + (1 | plot) + (1 | month), data=frl[depth=="5"]), ddf="Kenward-Roger")
anova(lmer(FRL ~ treatment * season + (1 | plot) + (1 | month), data=frl[depth=="5"]), ddf="lme4")

anova(lmer(FRL ~ treatment * season + (1 | plot), data=frl[depth=="5"]))
anova(lmer(FRL ~ treatment * season + (1 | plot), data=frl[depth=="5"]), ddf="Kenward-Roger")
anova(lmer(FRL ~ treatment * season + (1 | plot), data=frl[depth=="5"]), ddf="lme4")

anova(lmer(FRL ~ treatment * season + (1 | month), data=frl[depth=="5"]))
anova(lmer(FRL ~ treatment * season + (1 | month), data=frl[depth=="5"]), ddf="Kenward-Roger")
anova(lmer(FRL ~ treatment * season + (1 | month), data=frl[depth=="5"]), ddf="lme4")

```

#### were the duplicate frlJul columns not analyzed?

No.
```{r}
# remove the two July columns that I created above
FRL_cols_2 <- c("FRL5cmJun13",	"FRL5cmAug13",	"FRL5cmSep13",	"FRL5cmOct13",	"FRL5cmNov13",	"FRL5cmDec13",	"FRL5cmJan14",	"FRL5cmFeb14",	"FRL5cmMar14",	"FRL5cmApr14",	"FRL5cmMay14",	"FRL10cmJun13",	"FRL10cmAug13",	"FRL10cmSep13",	"FRL10cmOct13",	"FRL10cmNov13",	"FRL10cmDec13",	"FRL10cmJan14",	"FRL10cmFeb14",	"FRL10cmMar14",	"FRL10cmApr14",	"FRL10cmMay14")
frl2 <- melt(root, id.vars=c('treatment', 'plot'), measure.vars=FRL_cols_2, variable.name='label', value.name='FRL')
frl2[, treatment:=factor(treatment)] #CT-control; LA-litter addition; LR-litter removal
frl2[, label:=as.character(label)]
frl2[, depth:=factor(ifelse(substr(label, 4, 4)=='5', 5, 10))] # depth is 4th character in label
frl2[, month:=substr(label, nchar(label)-4, nchar(label)-2)] # month is in label
dry <- c('Jan', 'Feb', 'Mar', 'Apr') # wet is everything else
# May-Dec is "wet" per fig S1
frl2[, season:=factor(ifelse(month %in% dry, 'dry', 'wet'))]

anova(lmer(FRL ~ treatment * season + (1 | plot) + (1 | month), data=frl2[depth=="5"]))

```

#### is my dry season coded correctly?
yes

```{r}
unique(frl[season=="dry", month])
```

### Figure 3

```{r fig3-repro}
# fine root length 5 cm
frl5_se1 <- frl[depth==5,.(N=.N,
              FRL=mean(FRL),
              SE=sd(FRL)/sqrt(.N)),
              by=.(treatment, season)]
frl5.p1 <- frl[depth==5, .(FRL=mean(FRL)),
              by=.(treatment, season, plot)]
frl5_se2 <- frl5.p1[,.(N=.N,
              FRL=mean(FRL),
              SE=sd(FRL)/sqrt(.N)),
              by=.(treatment, season)]
# re-order factors
frl5_se2[, season:=factor(season, c("wet", "dry"))]
frl5_se2[, treatment:=factor(treatment, c("LR", "CT", "LA"))]
pd <- position_dodge(.6)
gg1 <- ggplot(data=frl5_se2, aes(x=season, y=FRL, fill=treatment)) +
  geom_col(position=pd, width=0.6, color='black') +
  geom_errorbar(aes(ymin=FRL-SE, ymax=FRL+SE), width=.2, position=pd) +
  ylab("FRL (0-5cm)") +
  scale_fill_manual(values=c("white", "gray", "black")) +
  theme_minimal(base_size=14)

# fine root length 10 cm
frl10.p1 <- frl[depth==10, .(FRL=mean(FRL)),
              by=.(treatment, season, plot)]
frl10_se2 <- frl10.p1[,.(N=.N,
              FRL=mean(FRL),
              SE=sd(FRL)/sqrt(.N)),
              by=.(treatment, season)]
# re-order factors
frl10_se2[, season:=factor(season, c("wet", "dry"))]
frl10_se2[, treatment:=factor(treatment, c("LR", "CT", "LA"))]
pd <- position_dodge(.6)
gg2 <- ggplot(data=frl10_se2, aes(x=season, y=FRL, fill=treatment)) +
  geom_col(position=pd, width=0.6, color='black') +
  geom_errorbar(aes(ymin=FRL-SE, ymax=FRL+SE), width=.2, position=pd) +
  ylab("FRL (0-10cm)") +
  scale_fill_manual(values=c("white", "gray", "black")) +
  theme_minimal(base_size=14)

# fine root length production
frl_prod.p1 <- frl_prod[, .(FRL_prod=mean(FRL_prod)),
              by=.(treatment, season, plot)]
frl_prod_se2 <- frl_prod.p1[,.(N=.N,
              FRL_prod=mean(FRL_prod),
              SE=sd(FRL_prod)/sqrt(.N)),
              by=.(treatment, season)]
# re-order factors
frl_prod_se2[, season:=factor(season, c("wet", "dry"))]
frl_prod_se2[, treatment:=factor(treatment, c("LR", "CT", "LA"))]
pd <- position_dodge(.6)
gg3 <- ggplot(data=frl_prod_se2, aes(x=season, y=FRL_prod, fill=treatment)) +
  geom_col(position=pd, width=0.6, color='black') +
  geom_errorbar(aes(ymin=FRL_prod-SE, ymax=FRL_prod+SE), width=.2, position=pd) +
  ylab("FRL production") +
  scale_fill_manual(values=c("white", "gray", "black")) +
  theme_minimal(base_size=14)

# fine root length production
frl_surv.p1 <- frl_surv[, .(FRL_surv=mean(FRL_surv)),
              by=.(treatment, season, plot)]
frl_surv_se2 <- frl_surv.p1[,.(N=.N,
              FRL_surv=mean(FRL_surv),
              SE=sd(FRL_surv)/sqrt(.N)),
              by=.(treatment, season)]
# re-order factors
frl_surv_se2[, season:=factor(season, c("wet", "dry"))]
frl_surv_se2[, treatment:=factor(treatment, c("LR", "CT", "LA"))]
pd <- position_dodge(.6)
gg4 <- ggplot(data=frl_surv_se2, aes(x=season, y=FRL_surv, fill=treatment)) +
  geom_col(position=pd, width=0.6, color='black') +
  geom_errorbar(aes(ymin=FRL_surv-SE, ymax=FRL_surv+SE), width=.2, position=pd) +
  ylab("FRL survival") +
  scale_fill_manual(values=c("white", "gray", "black")) +
  theme_minimal(base_size=14)
gg1
gg2
gg3
gg4
```

```{r fig3, echo=FALSE}
fn <- "fig3.png"
if(knitr_path==TRUE){
  image_path <- paste("../images", folder, fn, sep="/")
}else{
  image_path <- paste("/images", folder, fn, sep="/")
}
knitr::include_graphics(image_path)
```

The means look the same but my SE of the means are quite different than the reported SEMs. I used #2 (SEs computed from plot means averaged over months) although I would prefer using the modeled SEs from the fit model.

#### Could the SEs be #2 but among months averaged over plots?
```{r}
frl5.p1 <- frl[depth==5, .(FRL=mean(FRL)),
              by=.(treatment, season, month)]
frl5_se2.2 <- frl5.p1[,.(N=.N,
              FRL=mean(FRL),
              SE=sd(FRL)/sqrt(.N)),
              by=.(treatment, season)]
frl5_se2
frl5_se2.2
```


### post hoc tests for fig. 3
fine root length 0-5cm
```{r}
frl5.emm <- emmeans(frl5.fit, specs=c("treatment", "season"))
contrast(frl5.emm, method="revpairwise", simple = "each", combine=FALSE)$"simple contrasts for treatment"
```

fine root length 0-10 cm
```{r}
frl10.emm <- emmeans(frl10.fit, specs=c("treatment", "season"))
contrast(frl10.emm, method="revpairwise", simple = "each", combine=FALSE)$"simple contrasts for treatment"

```

fine root length production
```{r}
frl_prod.emm <- emmeans(frl_prod.fit, specs=c("treatment", "season"))
contrast(frl_prod.emm, method="revpairwise", simple = "each", combine=FALSE)$"simple contrasts for treatment"

```

fine root length survival
```{r}
frl_surv.emm <- emmeans(frl_surv.fit, specs=c("treatment", "season"))
contrast(frl_surv.emm, method="revpairwise", simple = "each", combine=FALSE)$"simple contrasts for treatment"

```
# Statistical red-flags

